### 2020-05-18 lineage_probabilities/lineage_probs.tex
draft of information about N(t) and lineage probs

Many popular approaches in population genetics use the distribution of coalescence rates between 
pairs of lineages in one or more populations to infer effective population sizes over time 
\cite{li2011inference,sheehan2013estimating,schiffels2014inferring}
or split times 
and subsequent migration rates between populations 
\cite{wang2020tracking}.
Given a demographic model, the 
\texttt{msprime} DemographyDebugger is able to compute coalescence rates for two or more lineages drawn 
from one or more populations at specified times in the past, providing a ground truth for coalescent rate-based 
inference methods. To compute coalescence rates for a set of times $Ts=(t_0, t_1, \ldots)$ in the 
past, we write
\begin{lstlisting}[frame=single]
dbg = demography.debug()
rates = dbg.coalescence_rate_trajectory(Ts, num_samples)
\end{lstlisting}
where \texttt{num_samples} is a list with length equal to the number of populations that
specifies the number of sampled lineages from each population. For example, 
to get pairwise rates within the same 
population, this list would contain zeros for all populations except for the population index of interest, 
which would have a $2$. To get the rate of coalescence for two lineages sampled from different 
populations, we would set \texttt{num_samples} to have ones in those population indices.

More simply, \texttt{msprime} can also compute the probability of a lineage that was in a given
population at some time to be found in other populations (through continuous or mass migration events). 
For a given sample time $t\geq0$ and any list of times $Ts$ in the past, we write
\begin{lstlisting}[frame=single]
prob = dbg.lineage_probabilities(Ts, sample_time=t)
\end{lstlisting}
This function is also used to determine the time intervals that lineages can be found in each population,
given a list of contemporary and/or ancient samples. For a given sampling configuration, by writing
\begin{lstlisting}[frame=single]
lineage_dict = dbg.possible_lineage_locations(samples=samples)
\end{lstlisting}
we get a dictionary with epoch intervals as keys whose values are a list that indicates True or False for 
whether each population could contain ancestral material to our samples.
If no sampling configuration is provided, we assume we sample lineages from each population at time zero.