### 2020-05-21 msprimepaper.tex
latex source file

\documentclass{article}
\usepackage[round]{natbib}
\usepackage[english]{babel}%
\usepackage[T1]{fontenc}%
\usepackage[utf8]{inputenc}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{geometry}%
\usepackage{dsfont}
\usepackage{verbatim}%
\usepackage{environ}%
\usepackage[right]{lineno}%
\usepackage{showkeys}
%%
% local definitions
\newcommand{\msprime}[0]{{\texttt{msprime} }}
\newcommand{\tskit}[0]{{\texttt{tskit} }}
\newcommand{\ms}[0]{{\texttt{ms} }}
\newcommand{\scrm}[0]{{\texttt{scrm} }}
\newcommand{\stdpopsim}[0]{{\texttt{stdpopsim} }}
 \newcommand{\be}{\begin{equation}}
 \newcommand{\ee}{\end{equation}}
 \newcommand{\bd}{\begin{displaymath}}
 \newcommand{\ed}{\end{displaymath}}
\newcommand{\IN}{\ensuremath{\mathds{N}}}%
\newcommand{\EE}[1]{\ensuremath{\mathds{E}\left[ #1 \right]}}%
\newcommand{\one}[1]{\ensuremath{\mathds{1}_{\left\{ #1 \right\}}}}%
\newcommand{\prb}[1]{\ensuremath{\mathds{P}\left( #1 \right) } }%

\NewEnviron{esplit}[1]{%
\begin{equation}
\label{#1}
\begin{split}
  \BODY
\end{split}\end{equation}
}

\begin{document}%

\linenumbers

\subsection*{Multiple merger coalescents}%
\label{mmc}%

Kingman's coalescent assumes that at most two ancestral lineages can
merge at each merger event.  This follows from the fact that the
Kingman coalescent can be shown to be the process describing the
random ancestral relations of gene copies sampled from a population
evolving according to the Wright-Fisher model (or similar models) of
genetic evolution.  Multiple-merger coalescent models can also be
obtained from appropriate extensions of the Moran
model \citep{EW06,HM12}, i.e.\ where a single randomly picked individual
(e.g.\ in a haploid population), or a pair of diploid parent
individuals \citep{BBE13}, produces offspring in each reproduction
event.  Convergence to the Kingman coalescent follows from certain
assumptions on the offspring distribution \citep{S99,MS01,SW08}.  These
assumptions may be violated in certain highly fecund
organisms \citep{hedgecock_94,B94,HP11,A04,irwin16}.  We take `high
fecundity' to mean the ability of individuals to produce numbers of
offspring on the order of the population size.  A key issue regarding
high fecundity involves the concept of `sweepstakes
reproduction' \citep{hedgecock_94}, where a few individuals may produce
the bulk of the offspring in any given generation.  Sweepstakes
reproduction is not captured by the Wrigth-Fisher model (or similar
models).  Population models, which do capture high fecundity and
sweepstakes reproduction, are in the domain of attraction of
`multiple-merger' coalescents \citep{DK99,P99,S99,S00}, in which a
random number of ancestral lineages may merge at any given time.  The
term `multiple merger' refers to a merger of at least three ancestral
lineages, or the simultaneous \citep{S00,MS01} merger of at least four
lineages in at least two distinct groups, and each group involving at
least two lineages.  Multiple-merger coalescent processes have also
been shown to be relevant for modeling the effects of selection on
gene genealogies \citep{Gillespie909,DS04}.

Multiple-merger coalescents, in which a random number of ancestral
lineages may merge at a given time in one group, and only one such
group of lineages can merge at any given time (asynchronous multiple
mergers), are referred to as $\Lambda$-coalescents; the rate at which  a given group of $k$ lineages out of a total of  $b$ lineages merges  is given by ($a\ge  0$ constant)
\begin{equation}\label{lambdabk}
\lambda_{b, k} =  \int_0^1  x^{k-2}(1-x)^{b-k}\Lambda(dx) + a\one{k=2}, \quad 2 \le k \le b, 
\end{equation}
where $\Lambda$ is a finite measure on the  (Borel subsets of)  unit interval without an atom at zero \citep{DK99,P99,S99}.    The  total rate  is given by
\be\label{lambdab}
 \lambda_{b} = \int_0^1 \left(1 - (1-x)^{b} - bx(1-x)^{b-1} \right)x^{-2}\Lambda(dx) + \binom{b}{2},
\ee
\citep{S99}  which can be useful in applications, in particular  provided the  antiderivative  can be explicitly identified.

A larger class of multiple-merger coalescents involving simultaneous
multiple mergers of distinct groups of ancestral lineages also exists
\citep{S00}. These are commonly referred to as $\Xi$-coalescents, and
can be shown to be the limits of ancestral processes derived from
population models incorporating diploidy (or more general polyploidy)
\citep{BBE13,Blath2016}, and certain models of selection \citep{DS04}.
To describe a general $\Xi$-coalescent let $\Delta$ denote the
infinite simplex \be\label{Delta} \Delta := \{ (x_1, \ldots ): x_1 \ge
x_2 \ge \cdots \ge 0, \sum_{j}x_j \le 1\}; \ee for any
$r \in \{1,2, \ldots\}$ let $k_1, \cdots, k_r \ge 2$, and
$b = s + k_1 + \cdots + k_r$ be the total number of blocks in a given
partition ($s = b - k_1 - \cdots - k_r$ is the number of blocks not
participating in the mergers at the given time).  The existence of
simultaneous multiple-merger coalescents was proved by \cite{S00}.
There exists a finite measure $\Xi$ on $\Delta$, with
$\Xi = \Xi_0 + a\delta_0$, the measure $\Xi_0$ has no atom at zero,
$a >0$ is fixed, and the rate at which one sees a given (simultaneous)
merger of ancestral lineages with merger sizes $k_1, \ldots, k_r$,
$s = n - k_1 - \cdots - k_r$, is given by
\begin{esplit}{xi}
  \lambda_{n; k_1, \ldots, k_r; s}  & = \int_\Delta  \sum_{\ell = 0}^s \sum_{\substack {i_1, \ldots, i_{r+\ell} = 1\\ \text{all distinct}} }^\infty  \binom{s}{\ell} x_{i_1}^{k_1} \cdots  x
_{i_{r}}^{k_r} x_{i_{r+1}} \cdots x_{i_{r+\ell}}\left(1 - \sum_{j=1}^\infty x_j \right)^{s-\ell} \frac{1}{ \sum_{j=1}^\infty x_j^2 } \Xi_0(dx)   \\
  & +  a\one{r=1, k_1 = 2}.
\end{esplit}%
The number of such $(k_1, \ldots, k_r)$ mergers is
\be\label{N}
    \mathcal{N}(b; k_1, \ldots, k_r ) = \binom{b}{k_1 \ldots k_r\, s} \frac{1}{ \prod_{j=2}^b\ell_j!  },
\ee
\citep{S00},   in particular  $\mathcal{N}(b;2) = b(b-1)/2$, and one can compute the total rate of a  $(k_1, \ldots, k_r)$ merger as
\be\label{lambdabkall}
      \lambda(n; k_1, \ldots, k_r)         =    \mathcal{N}(b; k_1, \ldots, k_r ) \lambda_{n; k_1, \ldots, k_r; s}.
\ee

The total rate is given by, with 
$n\ge 2$ denoting the total  number of ancestral lineages,  
\be
  \label{Xilambdab}
  \lambda_{n} = \int_\Delta \left(1 - \sum_{\ell = 0}^n \sum_{i_1 \neq
  \cdots \neq i_\ell } \binom{n}{\ell} x_{i_1}\cdots x_{i_\ell}\left(1
  - \sum_{i=1}^\infty x_i\right)^{n-\ell} \right)
  \frac{1}{\sum_{j=1}^\infty x_j^2}\Xi_0(dx) + a\binom{n}{2} \ee
  \citep{S00}.  Viewing coalescent processes strictly as mathematical
  objects, it is clear that the class of $\Xi$-coalescents contains
  $\Lambda$-coalescents as a specific example (i.e.\ allowing at most
  one group of lineages to merge each time), and the class of
  $\Lambda$-coalescents contain the Kingman-coalescent as a special
  case.  However, viewed as limits of ancestral processes derived from
  specific population models they are not nested, since one would
  obtain $\Lambda$-coalescents when deriving coalescent processes from
  haploid population models incorporating sweepstakes reproduction and
  high fecundity, and $\Xi$-coalescents for diploid populations.  One
  should therefore apply the models as appropriate, i.e.\
  $\Lambda$-coalescents to data (e.g.\ mtDNA data) inherited in a
  haploid fashion, and $\Xi$-coalescents to e.g.\ autosomal data
  inherited in diploid or polyploid fashion \citep{Blath2016}.

In \msprime we have incorporated two examples of multiple-merger
coalescents.  One is a diploid extension \citep{BBE13} of the haploid
model of sweepstakes reproduction considered by \cite{EW06}, which is
a haploid Moran model adapted to sweepstakes reproduction.  Let $N$
denote population size, and take $\psi \in (0,1]$ to be fixed.  In
every generation, with probability $1-\varepsilon_N$ a single
individual (picked uniformly at random) perishes, and one of the
surviving individuals (sampled uniformly at random) produces one
offspring; with probability $\varepsilon_N$ a total of
$\lfloor \psi N \rfloor$ individuals perish, and of the remaining
individuals a single individual produces $\lfloor \psi N \rfloor -1 $
offspring.  Taking $\varepsilon_N = 1/N^\gamma$ for some $\gamma > 0$,
\cite{EW06} obtain specific examples of $\Lambda$-coalescents, where
the $\Lambda$ measure in Eq \eqref{lambdabk} is a point mass at
$\psi$.  The simplicity of this model does allow one to obtain some
explicit mathematical results (see e.g.\
\cite{EF2018,Matuszewski2017,Der2012,Freund2020}). The model
considered by \cite{EW06} has also been applied in algorithms for
simulating gene genealogies within phylogenies \citep{Zhu2015}. The
 specific model incorporated into \msprime is the diploid version
\citep{BBE13} of the model studied by \cite{EW06}, which would be
necessary in order to incorporate recombination.  In the model
considered by \cite{BBE13}, a single pair of diploid individuals
contribute offspring in each generation, selfing is excluded, and each
offspring is assigned one chromosome from each parent. There are,
therefore, four parent chromosomes involved in each reproduction
event, which can lead to up to four simultaneous mergers.  Let
\begin{esplit}{Cconst}
C_{b; k_1, \ldots, k_r;s } &=  \frac{4}{\psi^2} \sum_{\ell = 0}^{s \wedge (4-r)} \binom{s}{\ell} (4)_{r+\ell} (1-\psi)^{s-\ell }\left( \frac{\psi}{4} \right) ^{k_1 + \cdots + k_r + \ell},  \\
\end{esplit}
where  $C_{b; k_1, \ldots, k_r;s }$ corresponds to the coalescence rate $ \lambda_{b;k_1, \ldots, k_r;s}$ (see Eq. \eqref{xi}) of a  $\Xi$-coalescent on  $(\psi/4, \psi/4, \psi/4, \psi/4, 0, \ldots)$.
The total merger rate (see Eq.\ \eqref{lambdabkall}) is then 
\be\label{xidirlambdabk}
      \lambda(b;k_1, \ldots, k_r) =    \mathcal{N}(b; k_1, \ldots, k_r ) \left( \frac{c\psi^2/4}{1 +  c\psi^2/4}C_{b; k_1, \ldots, k_r;s } +     \frac{ \one{r=1, k_1 = 2} }{1 +  c\psi^2/4}  \right),
\ee
and the total coalescence rate (see Eq.\ \eqref{xilambdab}) becomes
    \begin{esplit}{xilambdab}
\lambda_b & =  \frac{c\psi^2/4}{1 +  c\psi^2/4}  \frac{4 }{\psi^2 } \left(1 - \sum_{\ell = 0 }^{b\wedge 4} \binom{b}{\ell} (4)_\ell \left( \frac{\psi}{4} \right)^\ell (1 - \psi)^{b-\ell } \right) +    \frac{1}{1 +  c\psi^2/4} \binom{b}{2}.  \\
\end{esplit}
The interpretation of Eq.\ \eqref{xilambdab} is that with probability  $1/(1 + c\psi^2/4)$  a `small' reproduction event occurs, in which the parent pair produce one  diploid offspring; with probability  $c\psi^2/4/(1 + c\psi^2/4)$ a `large' reproduction event occurs, in which  the parent pair produce  $\lfloor \psi N \rfloor$ offspring.

The other multiple-merger coalescent model incorporated in \msprime is
derived from an adaptation of the haploid population model considered
by \cite{schweinsberg03} to diploid populations \citep{BLS15}.  In the
haploid version, in each generation individuals independently produce
random numbers of juveniles, where the random  number of juveniles $(X)$ produced by
a given individual  has the stable law \be\label{jX} \lim_{k\to
\infty} C k^\alpha \prb{X \ge k} = 1 \ee with index $\alpha > 0$, and
$C > 0$ is a normalising constant.   One can interpret  Eq.\ \eqref{jX} as  stating the form of the  probability distribution  for number of juveniles at least on the order of the population size.     If the random  total number of juveniles $(S_N)$ produced in this way is at least the population size $(N)$, then one samples  $N$ juveniles uniformly at random without replacement to form the next generation of  reproducing individuals; if  $S_N < N$ one simply  carries on with  the same  set of individuals.   However,   assuming  $\EE{X} > 1$, one can show that $\prb{S_N < N}$ decays exponentially fast in $N$ \citep{schweinsberg03}.         If $\alpha \ge 2$ the ancestral
process converges to the Kingman-coalescent; if $1 \le \alpha < 2$ the
ancestral process converges to a specific case of a
$\Lambda$-coalescent, where the $\Lambda$ measure in Eq.\
\eqref{lambdabk} is associated with the Beta$(2-\alpha, \alpha)$
probability distribution, i.e.\
\be\label{Fbeta}
    \Lambda(dx) = \one{0< x \le 1} \frac{1}{B(2-\alpha,\alpha)} x^{1 - \alpha}(1-x)^{\alpha - 1}  dx,
\ee
where $B(2-\alpha,\alpha)$ is the beta function $B(a,b) = \Gamma(a)\Gamma(b)/\Gamma(a+b)$, $a,b > 0$ \citep{schweinsberg03}.    This model has been adapted to diploid populations  by                      \cite{BLS15}, where  the resulting coalescent process  is a  four-fold $\Xi$-coalescent on  $(x/4, x/4, x/4, x/4, 0, 0, \ldots)$, where $x$ is a random variate with  the Beta$(2-\alpha,\alpha)$ distribution.  The merger rate (see Eq.\ \eqref{xi}) is then ($1 \le r \le 4$)
\be\label{xibeta}
   \lambda_{b;k_1, \ldots, k_r} = \sum_{\ell = 0}^{ (b - k)\wedge (4-r) } \binom{b-k}{\ell} \frac{ (4)_{r+\ell} }{4^{k+\ell}} \frac{B(k+\ell - \alpha, b-k-\ell + \alpha ) }{B(2-\alpha,\alpha)}
\ee
\citep{Blath2016,BLS15}. The interpretation of Eq.\ \eqref{xibeta} is  that the random   number of diploid  juveniles each  diploid pair of parents  produces  is  governed by  the law in Eq.\ \eqref{jX},   and   each diploid juvenile  is assigned  one chromosome  from each parent (selfing is excluded).

The model in Eq.\ \eqref{jX} assumes that individuals can produce
arbitrarily large numbers of juveniles. Considering diploid juveniles,
this assumption is probably rather strong, since diploid juveniles are
at least fertilised eggs, and so it is reasonable to suppose that the
number of juveniles surviving to the  particular life stage we are modeling  cannot be
arbitrarily large.  With this in mind, we also consider an adaptation
of the Schweinsberg model, where the random number of juveniles $(X)$
produced by a   given parent pair  is distributed according to
\be\label{jtr}
  \prb{X=k} =   \one{1 \le k \le \phi(N)} \frac{\phi(N+1)^\alpha }{ \phi(N+1)^\alpha - 1 }  \left( \frac{1}{k^\alpha} - \frac{1}{(k+1)^\alpha}  \right) ,
\ee
where $\phi(N)$ is a deterministic strict upper bound on the number of juveniles produced by  any given parent pair (see also \citep{Eldon2018}).   One can follow the calculations in  \citep{schweinsberg03} or \citep{BLS15}  to show  that  if $1 < \alpha < 2$   then  $(k = k_1 + \cdots + k_r)$ then the merger rate (see Eq.\ \eqref{xibeta}) is
\be
   \lambda_{b;k_1, \ldots, k_r} =  \sum_{\ell = 0}^{ (b - k)\wedge (4-r) } \binom{b-k}{\ell} \frac{ (4)_{r+\ell} }{4^{k+\ell}} \frac{B(M; k+\ell - \alpha, b-k-\ell + \alpha ) }{B(M;2-\alpha,\alpha)}
\ee
with $B(z;a,b) := \int_0^z t^{a-1}(1-t)^{b-1}dt$ for  $a,b>0$ and $0< z\le 1$, and
\be
M :=  \frac{K}{K+m} \one{\phi(N) = KN} + \one{\phi(N)/N \to \infty }
\ee
where $K > 0$ is a constant and  $m := \lim_{N\to \infty} \EE{X} = 1 + 2^{1-\alpha}/(\alpha - 1)$ \citep{CDEE2020,AEKKZ2020}.    In other words,  the measure $(\Lambda)$ driving the multiple mergers is of the same form as in Eq.\ \eqref{Fbeta}  with $0 < x \le M$ in the case $1 < \alpha < 2$ and $\phi(N) \ge KN$.   If $\alpha \ge 2$ or $\phi(N)/N \to 0$ then  the  ancestral process converges (in the sense of convergence of  finite-dimensional distributions) to  the Kingman-coalescent \citep{CDEE2020,AEKKZ2020}.

\cite{Becheler2020}  investigate a general framework for  constructing efficient algorithms for general  coalescent processes.

%assumption can be
%violated in some situations. Lambda coalescents
%%~\citep{donnelly1999particle,schweinsberg2000coalescents}
%%HybridLambda~\citep{zhu2015hybrid}

\begin{comment}%
A coalescent process is a continuous-time Markov process taking values among the
partitions of $\IN := \{1,2, \ldots \}$, such that the restriction to
any finite $n \in \IN $ takes values among partitions of
$[n] := \{1, 2, \ldots, n\}$.  Write $\one{A} = 1$ if $A$ holds, and zero otherwise.   Let $\cP_n$ denote the set of
partitions of $[n]$.  In the classical Kingman-coalescent, the only
possible transitions are the mergers of pairs of blocks (elements of a
partition $\pi \in \cP_n$), one pair at a time.  The $n$ leaves
(corresponding to the sampled DNA sequences) are arbitrarily labelled
from 1 to $n$, and  the blocks of a partition represent the common ancestors
of the labels of each block.    The initial state is  (usually) taken as 
$\{ \{1\}, \ldots, \{n\}\}$, and the final state, i.e.\ when the most 
recent common ancester is reached, as $\{ [n]\}$. A block in a partition of
$[n]$ represents an ancestor of the leaves in the block, i.e.\ the block
$\{i_1, \ldots, i_k\}$ in a given partition of $[n]$ is an ancestor of the
$k$ leaves $i_1, \ldots, i_k \in [n]$, and the leaves  correspond to
arbitrarily labelled DNA sequences in the sample.  
\end{comment}

\bibliographystyle{plainnat}
\bibliography{refs}

\end{document}%