### 2021-08-08 paper.tex

\documentclass{article}
\usepackage[round]{natbib}
\usepackage{listings}
\usepackage[english]{babel}%
\usepackage[T1]{fontenc}%
\usepackage[utf8]{inputenc}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{geometry}%
\usepackage{color}
\usepackage{graphicx}
\usepackage{dsfont}
\usepackage{verbatim}%
\usepackage{environ}%
\usepackage[right]{lineno}%
\usepackage{nameref}
\usepackage{tikz}
%\usepackage{showkeys}

% local definitions
\newcommand{\msprime}[0]{\texttt{msprime}}
\newcommand{\tskit}[0]{\texttt{tskit}}
\newcommand{\ms}[0]{\texttt{ms}}
\newcommand{\msms}[0]{\texttt{msms}}
\newcommand{\cosiTwo}[0]{\texttt{cosi2}}
\newcommand{\msHOT}[0]{\texttt{msHOT}}
\newcommand{\scrm}[0]{\texttt{scrm}}
\newcommand{\SLiM}[0]{\texttt{SLiM}}
\newcommand{\fwdpy}[0]{\texttt{fwdpy11}}
\newcommand{\stdpopsim}[0]{\texttt{stdpopsim}}
\newcommand{\SeqGen}[0]{\texttt{Seq-Gen}}
\newcommand{\Pyvolve}[0]{\texttt{Pyvolve}}
\newcommand{\phastsim}[0]{\texttt{phastsim}}
\newcommand{\discoal}[0]{\texttt{discoal}}
\newcommand{\ARGON}[0]{\texttt{ARGON}}

\newcommand{\jkcomment}[1]{\textcolor{red}{#1}}
\newcommand{\plrcomment}[1]{\textcolor{cyan}{#1}}
\newcommand{\grgcomment}[1]{\textcolor{yellow!60!red}{#1}}

\begin{document}

\title{Msprime 1.0: an efficient and extensible coalescent simulation framework}
\author{Author list to be filled in
}
% \address{

%%Affiliations:
% Franz Baumdicker:
% Cluster of Excellence "Controlling Microbes to Fight Infections", Mathematical and Computational Population Genetics, University of Tübingen, Germany

% Consuelo D. Quinto-Cortés
%National Laboratory of Genomics for Biodiversity (LANGEBIO), Unit of Advanced Genomics, CINVESTAV, Irapuato, Mexico

% Peter L. Ralph
% Institute of Ecology and Evolution, Departments of Mathematics and Biology, University of Oregon, USA

% \author[n]{Graham Gower}
% \affil[n]{Lundbeck GeoGenetics Centre, Globe Institute, Faculty of Health and Medical Sciences, University of Copenhagen, Denmark}

% \section*{Contact:} \href{jerome.kelleher@bdi.ox.ac.uk}{jerome.kelleher@bdi.ox.ac.uk}

\maketitle

\begin{abstract}
Coalescent simulation is a key tool in population genetics and
has been integral to theoretical developments since its beginnings.
Because of the ease with which the basic coalescent model can be simulated,
a large number of different simulators have been developed. However,
the coalescent with recombination is far more challenging to simulate,
and, until recently, it was not possible to simulate efficiently.
The \msprime\ software has revolutionised
coalescent simulation, making it possible to simulate millions
of whole genomes for the first time. We summarise the many features
of \msprime\ 1.0, which is built around a core model of efficiently
implementing recombination via the succinct tree sequence data
structure. Moreover, development of \msprime\ is collaborative
and open source, which reduces duplication of effort and promotes
software quality via community engagement.
% "Bazaar" development model: https://en.wikipedia.org/wiki/The_Cathedral_and_the_Bazaar
\end{abstract}

\textbf{Keywords:} Coalescent simulation, Python

\section*{Introduction}

The coalescent
process~\citep{kingman1982coalescent,hudson1983testing,tajima1983evolutionary}
models the ancestry of a set of sampled genomes. Specifically,
the coalescent provides a mathematical description of the
genealogical tree that relates sampled genomes to one another.
The coalescent has proved to be a powerful model,
and it is now central to
population genetics~\citep{hudson1990gene,hein2004gene,wakely2008coalescent}.
Among the most impactful contributions of the coalescent since its origins
~\citep{hudson1983testing} is that it provides a path towards more efficient
simulation of population genetic samples than is possible using traditional,
forwards in time simulation.
As a result of this great utility, dozens of different simulation tools have been
developed over the decades~\citep{carvajal2008simulation,liu2008survey,
arenas2012simulation,yuan2012overview,hoban2012computer,yang2014critical,
peng2015genetic}.
Two technological developments of recent years, however,
pose major challenges to the majority of these methods.
% Some citations here? Darwin tree of life maybe?
Firstly, whole genome sequencing---and in particular fourth-generation
sequencing technologies---have made chromosome-level assemblies
commonplace. Thus, modelling genetic variation data as a series of unlinked loci,
a previously reasonable assumption, is no longer defensible and
we must fully account for recombination.
While a genealogical tree relating $n$ samples in the single-locus coalescent
can be simulated in $O(n)$ time~\citep{hudson1990gene},
the coalescent with recombination, which generates multiple correlated trees, is
far more challenging.
% JK will need to revise this sentence, in light of what we'll be saying
% down further about needing the SMC for very large rho.
Programs such as Hudson's classical \ms~\citep{hudson2002generating}
can only simulate short segments under the influence of recombination,
and, until recently, approximate models of
recombination~\citep{mcvean2005approximating,staab2015scrm}
were necessary to simulate whole chromosomes.

The second major challenge to existing coalescent simulators is the
exponential increase in sample size of genetic studies. Human datasets like the
UK Biobank~\citep{bycroft2018genome} and
gnomAD~\citep{karczewski2020mutational} now consist of hundreds of
thousands of genomes and many other datasets on a similar scale
are becoming available~\citep{tanjo2021practical}.
Classical simulators such as \ms\ and even fast approximate methods
such as \scrm\ simply cannot cope with such a large number of samples.

The \msprime\ simulator~\citep{kelleher2016efficient,kelleher2020coalescent}
has greatly increased the scope of coalescent simulations,
and it is now straightforward to simulate millions of whole chromosomes.
The ``succinct tree sequence'' data
structure~\citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring,
wohns2021unified},
introduced as part of \msprime\, makes it possible to store
such huge simulations in a few gigabytes, several orders
of magnitude smaller than commonly used formats.
The succinct tree sequence has also lead to major advances in forwards-time
simulation~\citep{kelleher2018efficient,haller2018tree},
ancestry inference~\citep{kelleher2019inferring,wohns2021unified}
and calculation of population genetic
statistics~\citep{kelleher2016efficient,ralph2020efficiently}.
Through a rigorous open-source community development process,
\msprime\ has gained a large number of features since its introduction,
making it a highly efficient and flexible platform for population
genetic simulation.
This paper marks the release of \msprime\ 1.0, a state-of-the-art
coalescent simulator both with respect to efficiency as well as feature-richness.
Here we provide an overview of the extensive abilities of \msprime\ 1.0,
demonstrate its performance advantages over alternative software,
and finally lay out future plans for
further open-source development of what we hope will become an industry
standard piece of computational infrastructure for the genetics community.

\section*{Results}

\jkcomment{Give a quick roadmap,
like, we discuss the underlying tree sequence object,
then mutations (because it's quite separate from the rest)
and then a bunch of new features for ancestry simulation. There's no
point in listing it out section by section.}

% JK: commented out this text as it's misleading and would confuse
% people. We should adapt it when writing the roadmap when the structure
% has settled down.

% In this section we describe the main features of \msprime\ 1.0. Methodologically,
% there exists a logical separation between simulating coalescent genealogies and
% adding neutral mutations onto these genealogies. Simulating genetic ancestry
% requires methods that model the underlying reproductive process (e.g. Moran,
% Wright-Fisher, etc.) as well as accounting for recombination of various sorts
% (including gene conversion), demography, and selection. These methods are
% discussed in the first N subsections below. One of the strengths of the
% approach taken in msprime is that neutral mutations can be retrospectively
% added to simulated genealogies. Methods to generate mutations, and thereby
% generate genetic variation, are therefore discussed in a separate and subsequent
% subsection. Finally, we describe the \msprime\ API, which is designed to enforce
% the distinction between simulating genealogies and mutations; discuss the
% interoperability provided by the use of \tskit\, the tree sequence toolkit;
% and highlight the open source development process which allows for \msprime\
% to be extended by addition of new features while ensuring robustness and
% maintainability of the core codebase.

% The data model
\subsection*{Tree sequences}
\label{sec-ts}

\begin{figure}
\begin{center}
\includegraphics[width=6cm]{illustrations/example_tree_sequence}
\end{center}
\caption{\label{fig-ts-example} Example tree sequence.
An example tree sequence describing genealogies and genotypes
for four samples at ten sites.
}
\end{figure}

One of the key reasons for \msprime's substantial performance advantage
over other simulators~\citep{kelleher2016efficient}
is its usage of the ``succinct tree sequence''
 data structure to represent simulation results.
The succinct tree sequence (usually abbreviated to ``tree sequence'')
is a data structure for concisely encoding
genetic ancestry and sequence variation.
The basic concepts were introduced by \cite{kelleher2016efficient}
and originally developed as part of \msprime. We subsequently
extracted the tree sequence code that was independent of simulations
from \msprime\ into
a standalone Python and C library, called \tskit~\citep{tskit2021tskit}.
While a full discussion of tree sequences and the capabilities
of \tskit\ is beyond the scope of this article, we summarise
some aspects that are important for simulation.

% Encoding ancestry
Given a set of sample chromosomes, the topology
of their genetic ancestry is described by a set of \emph{nodes}
and \emph{edges} (Fig.~\ref{fig-ts-example}).
A node in tree sequence terminology corresponds
to a single chromosome strand, associated with its birth time
and other optional information. A node may be
associated with an \emph{individual}, so that, for example, a diploid
individual is associated with two separate nodes (we omit
the individual table in Fig.~\ref{fig-ts-example} as it is optional).
Nodes define haplotypes that exist at a specific time within
some individual, and edges define genetic inheritance
between these nodes. An \emph{edge} consists of a
parent node, a child node and the left and right coordinates
of the contiguous genomic interval over which
the child inherited genetic material from the parent.
Critically, edges can span multiple marginal trees: for
example, in Fig.~\ref{fig-ts-example} the branch from node
0 to 4 is present in both marginal trees, and represented
by a single edge (the first row in the edge table).
This simple device, of explicitly associating tree nodes
with specific ancestral chromosomes and recording the
genomic intervals over which parent/child relationships exist,
generalises the original
``coalescence records'' concept~\citep{kelleher2016efficient},
and is the key to the efficiency of tree
sequences~\citep{
kelleher2018efficient,kelleher2019inferring,ralph2020efficiently}.

The tree sequence encoding of ancestry as a set of nodes and edges
as just described, is equivalent to a graph in which edges
are annotated with a set of disjoint genomic intervals.
This is closely related to the Ancestral Recombination Graph (ARG),
a term that is often loosely understood
to mean some representation of genetic
ancestry as a graph~\citep[e.g.][]{mathieson2020ancestry}.
In this sense a succinct tree sequence \emph{is} an ARG.
More technically, an ARG is a particular
representation of genetic ancestry as a graph based on the
flow of ancestral material through recombination and common
ancestor events~\citep{griffiths1991two,griffiths1997ancestral}.
In this sense, tree sequences can be used to \emph{encode}
ARGs without loss of information (although
the default minimal tree sequence encoding is
much more concise; see the \nameref{sec-arg} section).
It should be noted that, despite standardisation
efforts~\citep{cardona2008extended,mcgill2013graphml}, there is
no shared format for interchanging ARG data, and no supporting software
libraries.

% OK, back to tree sequences: what about sequence variation?
The final output of most population genetic simulations is some representation
of sequence variation among the specified samples. For coalescent
simulations, we usually have three steps: (1) simulate
the genetic ancestry, and optionally output the resulting marginal trees;
(2) simulate sequence evolution conditioned on this ancestry by generating
mutations (see the \nameref{sec-mutations} section); and (3) output the
resulting sequences by percolating the effects of the mutations through
the trees. Information about the mutations
themselves---e.g., where they have occurred on the
trees---is usually not retained or made available for subsequent analysis.
In \msprime, however, we skip step (3), instead using \tskit's
combined data model of ancestry and mutations
to represent the simulated sequences.
As illustrated in Fig.~\ref{fig-ts-example}, mutations are a
fully integrated part of \tskit's tree sequence data model, and genetic
variation is encoded by recording sites at which mutations
have occurred, and where each mutation at those sites
has occurred on the marginal tree.
Crucially, the genetic sequences themselves are never stored, or indeed directly
represented in memory (unless explicitly decoded using \tskit\ APIs).
It may at first seem inconvenient to have only this indirect representation
of the genome sequences, but it is extremely powerful.
Firstly, the storage space required for simulations is dramatically
reduced. For a simulation of $n$ samples with $m$ variant sites,
we would require $O(nm)$ space to store the sequence data as a variant
matrix. However, if this simulation was of a recombining genome
with $t$ trees, then the \tskit\ tree sequence encoding
requires $O(n + t + m)$ space,
assuming we have $O(1)$ mutations at each site~\citep{kelleher2016efficient}.
For large sample sizes, this difference is profound, making it
conceivable, for example, to store the genetic ancestry
and variation data for the entire human population on a
laptop~\citep{kelleher2019inferring}. As well as the huge
difference in storage efficiency, it is often far more efficient
to compute statistics of the sequence data from the
trees and mutations than it is to work with the sequences themselves.
For example, computing Tajima's $D$ from simulated data stored in the \tskit\ format
is several orders of magnitude faster than efficient variant matrix
libraries for large sample sizes~\citep{ralph2020efficiently}.

The vast genomic datasets produced during the SARS-CoV-2 pandemic
have highlighted the advantages of storing genetic variation data using the underlying trees.
\cite{turakhia2021ultrafast} propose the
Mutation Annotated Tree (MAT)
format (consisting of a Newick tree
and associated mutations in a binary format) and
the \texttt{matUtils} program as an efficient way to store
and process large viral datasets~\citep{mcbroome2021daily},
achieving excellent compression and processing performance.
Similarly, \phastsim~\citep{demaio2021phastsim}
was developed to simulate sequence evolution on such large SARS-CoV-2
phylogenies, and also outputs a Newick tree annotated with mutations (not
in MAT format) to avoid the bottleneck of generating
and storing the simulated sequences.
While these methods illustrate the advantages of the general
approach of storing ancestry and mutations rather than sequences,
they do not generalise beyond their immediate settings, and
no software library support is available.
% \msprime, on the other hand, provides a fast, powerful, and well-tested
% generic interface for generating and storing mutations on trees.

The software ecosystem built around \tskit\ is stable and mature;
simulators such as
\fwdpy~\citep{thornton2014cpp},
\SLiM~\citep{haller2019slim},
\stdpopsim~\citep{adrion2020community},
\texttt{Geonomics}~\citep{de2021geonomics}
and
\texttt{GSpace}~\citep{virgoulay2021gspace},
and inference methods such as
\texttt{tsinfer}~\citep{kelleher2019inferring},
\texttt{tsdate}~\citep{,wohns2021unified}
and \texttt{Relate}~\citep{speidel2019method}
use either the Python or C
APIs to support outputting results in tree sequence format.
Tree sequences are stored in an efficient binary file format, and
are fully portable across operating systems and processor
architectures. The \tskit\ library ensures interoperability
between programs by having strict definitions of how the
information in each of the tables is interpreted, and
stringent checks for the internal consistency of the data model.

% The input - how do we run simulations?
\subsection*{User interface}
\label{sec-sim-interface}

The majority of simulation packages are controlled either through
a command line interface~\citep[e.g.][]{hudson2002generating,kern2016discoal},
a text-based input file
format~\citep[e.g.][]{guillaume2006nemo,excoffier2011fastsimcoal,shlyakhter2014cosi2},
or a mixture of both.
Command line interfaces make it very easy to run simple
simulations, but as model complexity and the number of parameters specified increase,
they become difficult to understand and
error-prone~\citep{ragsdale2020lessons,gower2021demes}.
Specifying parameters through a text file alleviates this problem to a degree,
but lacks flexibility, for example, when running simulations with parameters
drawn from some distribution. In practice, for any reproducible simulation
project users will write a script
to generate the required command lines or input parameter files,
invoke the simulation engine, and process the results in some way.
This process is cumbersome and labour intensive, and
a number of packages have been developed
to allow simulations to be run directly in a high-level
scripting language~\citep{staab2016coala,parobek2017skelesim,gladstein2018simprily}.

The more recent trend has been to move away from this file and command-line
driven approach and to instead provide direct interfaces to the simulation
engines via an Application Programming Interface (API)~\citep[e.g.][]{
thornton2014cpp,kelleher2016efficient,becheler2019quetzal,haller2019slim}.
The primary interface for \msprime\ is through a thoroughly documented and
stable Python
API, which has encouraged the development of an ecosystem of downstream
% Found these through the Dependent packages on GitHub
tools~\citep{terhorst2017robust,chan2018likelihood,spence2019inference,
adrion2020community,adrion2020predicting, kamm2020efficiently,
mckenzie2020ipcoal, montinaro2020revisiting,
de2021geonomics,rivera2021simulation}.
As well as providing a stable and efficient platform for building
downstream applications, \msprime's Python API makes it much easier to
build reproducible simulation pipelines, as the entire workflow can
be encapsulated in a single script, and package and version
dependencies explicitly stated using the \texttt{pip}
or \texttt{conda} package managers.
For example, the errors made
in the influential simulation analysis of
\cite{martin2017human} were only detected because the pipeline
could be easily run and reanalysed~\citep{ragsdale2020lessons,martin2020erratum}.

A major change for the \msprime\ 1.0 release is the introduction of a new set of APIs,
designed in part to avoid sources of error (see the Demography section) but
also to provide more appropriate defaults while keeping compatibility with
existing code. In the new APIs, ancestry and mutation simulation are fully
separated, with the \texttt{sim\_ancestry} and \texttt{sim\_mutations}
functions replacing the legacy \texttt{simulate} function. Among other changes,
the new APIs default to discrete genome coordinates and finite sites mutations,
making the default settings more realistic and resolving a major source of confusion and error.
The previous APIs are fully
supported and tested, and will be maintained for the foreseeable future.
We also provide an \ms\ compatible
command line interface to support existing workflows. The
\texttt{msp} program has been extended to include new commands
for simulating ancestry and mutations separately. A particularly useful
feature is the ability to specify demographic models in
Demes format~\citep{gower2021demes} from the command line,
making simulation of complex demographies straightforward.

% The output - what do we get out?
\subsection*{Data analysis}
\label{sec-data-analysis}
The point of running simulations is to generate data, and there
are two important ways in which the \tskit\ tree sequences output by
\msprime\ are more useful than classical methods:
they are more efficient and they contain more information.
We will consider these points in turn.

The standard way of representing simulation data is
to render the results in a text format. For example,
\ms~\citep{hudson2002generating} outputs a set of sequences
and can also optionally output the marginal trees along
the genome in Newick format,
and variants of this approach are used by many simulators.
Text files have many advantages, but are slow to process
at scale.
The ability to efficiently process simulation results is
particularly important in simulation-based inference methods
such as Approximate Bayesian Computation
(ABC)~\citep{beaumont2002approximate,csillery2010approximate,wegmann2010abctoolbox}
and machine learning based
approaches~\citep{sheehan2016deep,chan2018likelihood,schrider2018supervised,
flagel2019unreasonable,sanchez2020deep}. Clearly, simulation efficiency is
crucial since the size and number of simulations that can be performed determines
the depth to which one can sample from the model and parameter space.
Equally important,
however, is the efficiency with which the simulation results can be
transformed into the specific input required by the inference method.
In the case of ABC, this is usually a set of summary statistics of the sequence
data, and methods avoid the bottleneck of parsing
text-based file formats to compute these statistics
by either developing their own
simulators~\citep[e.g.][]{cornuet2008inferring,lopes2009popabc}
or creating forked versions (i.e., modified copies) of existing
simulators~\cite[e.g.][]{thornton2006approximate,
hickerson2007msbayes,pavlidis2010msabc,huang2011mtml,quinto2018modeling},
tightly integrated with the inference method.
Modern approaches to ABC such as
ABC-RF~\citep{raynal2019abc,pudlo2016abc} and
ABC-NN~\citep{csillery2012abc,blum2010abc} use large
numbers of weakly informative statistics,
making the need to efficiently compute statistics from simulation
results all the more acute.

Classical text based output formats like \ms\ are inefficient to process,
but also lack a great deal of important information.
The tree-by-tree topology information output by simulators in Newick
format lacks any concept of node identity,
and means that we cannot reliably infer information about ancestors
from the output. Because Newick stores branch lengths rather
than node times, numerical precision issues also arise when summing
over many short branches~\citep{mcgill2013graphml}.
Numerous forks  of simulators have been created to access information not provided
in the output. For example, \ms\  has been forked to
output information about migrating segments~\citep{rosenzweig2016powerful},
information about ancestral lineages~\citep{chen2013asymptotic},
an array of population genetic statistics~\citep{ramos2007mlcoalsim}
% Any more?
and \ms's fork \msHOT~\citep{hellenthal2007mshot},
has in turn been forked to output information on local
ancestry~\citep{racimo2017archaic}.
All of this information
is either directly available by default in \msprime, or can be optionally
stored via options such as \texttt{record\_migrations} or
\texttt{record\_full\_arg} (see the \nameref{sec-arg} section) and
can be efficiently and conveniently processed via \tskit\ APIs.

These problems of inefficient and limited access to the simulated
data are comprehensively solved by \msprime\ and \tskit. Complete
information about ancestry and mutation in simulations is immediately
available to the user in a single integrated data structure
(rather than multiple independent files), with no intermediate
processing required.
Users can subsequently compute precisely the information that they
require using the powerful suite of population genetic statistics
and other methods available in \tskit.

% Stuff we simulate: mutations
\subsection*{Simulating mutations}
\label{sec-mutations}

% Explain the background: what have people done classically?
Because coalescent simulations are usually concerned with
neutral evolution (however, see the \nameref{sec-selection} section)
the problem of generating synthetic genetic variation can be decomposed into
two independent steps:
firstly, simulating genetic ancestry (the trees), then subsequently simulating
variation by superimposing mutation processes on those trees
(see Fig.~\ref{fig-mutated-trees}).
A number of programs exist to place mutations on trees: for instance,
the classical \SeqGen\ program~\citep{rambaut1997seq}
supports a range of different models of sequence evolution,
and various extensions to the basic
models have been proposed~\citep[e.g.][]{cartwright2005dna,fletcher2009indelible}.
Partly for efficiency and partly in the interest of
simplicity for users (i.e., to avoid intermediate text format conversions),
population genetic simulators have tended to
include their own implementations of mutation simulation, with
most supporting the infinite sites
model~\citep[e.g.][]{hudson2002generating}
but with several supporting a wide range of different models of sequence
evolution~\citep[e.g.][]{mailund2005coasim,excoffier2011fastsimcoal,
virgoulay2021gspace}. Thus, despite the logical separation between
the tasks of simulating ancestry and neutral sequence evolution,
the two have been conflated in practice.

\begin{figure}
    \includegraphics{illustrations/unmutated_tree}
    \includegraphics{illustrations/mutated_tree}
    \caption{
        \label{fig-mutated-trees}
        Visualisation of the separation between ancestry and mutation simulation:
        \texttt{sim\_ancestry} produces ancestry
        (\textbf{left;} shown are two trees along a 1Kb chunk of genome
        relating 200 samples),
        and \texttt{sim\_mutations} adds mutations
        (\textbf{right;} fourteen nucleotide mutations have been added).
        Graphics produced with \tskit's \texttt{draw\_svg} method.
        \plrcomment{TODO: fix up the axis labels, if we like this figure?}
        \jkcomment{Why 200 samples? Seems like far too many to me, how about 5?}
        \grgcomment{Add \texttt{sim\_ancestry} and \texttt{sim\_mutations} as subfigure titles?}
    }
\end{figure}

Part of the reason for this poor record of software reuse and modularity is the
lack of standardised file formats, and in particular, the absence of common
library infrastructure to abstract the details of interchanging simulation
data. Although \msprime\ also supports simulating both ancestry and mutations,
the two aspects are functionally independent within the software.
Both ancestry and mutation simulators are present in \msprime\
for reasons of convenience and history,
and could be split into separate packages.
The mature and efficient C and Python interfaces for
\tskit\ make it straightforward to add further information to
an existing file, and because of its efficient data interchange mechanisms,
there is no performance penalty for performing additional operations
in a different Python package.
Thanks to this interoperability, \msprime's
mutation generator can work with \emph{any} \tskit\ tree sequence,
be it simulated using \SLiM~\citep{haller2019slim} or
\fwdpy~\citep{thornton2014cpp}, or estimated from real
data~\citep{kelleher2019inferring,speidel2019method,wohns2021unified}.
It is a modular
component intended to fit into a larger software ecosystem, and
is in no way dependent on \msprime's ancestry simulator.

As well as providing a new API
that emphasises the logical split between ancestry and mutation simulation
(see the \nameref{sec-sim-interface} section),
we have greatly extended the sophistication of
\msprime's mutation generation engine for version 1.0,
achieving near feature-parity with \SeqGen.
We support a large number of mutation models, including the
JC69~\citep{jukes1969evolution},
F84~\citep{felsenstein1996hidden},
and GTR~\citep{tavare1986some} nucleotide models
and the BLOSUM62~\citep{henikoff1992amino}
and PAM~\citep{dayhoff1978} amino acid models.
Other models, such as the Kimura two and three
parameter models~\citep{kimura1980simple,kimura1981estimation},
can be defined easily and efficiently in
user code by specifying an arbitrary transition matrix between
any number of alleles (which can be arbitrary unicode strings).
Mutation rates can vary along the genome, and multiple mutation
models can be imposed on a tree sequence by overlaying mutations
in multiple passes.
We have extensively validated the results of mutation simulations
against both theoretical expectations and output from
\SeqGen~\citep{rambaut1997seq} and
\Pyvolve~\citep{spielman2015pyvolve}.

\begin{figure}
    \includegraphics[width=\textwidth]{figures/mutations-perf}
\caption{\label{fig-mutations-perf} Time required to run
\texttt{sim\_mutations} on tree sequences generated
by \texttt{sim\_ancestry} (with a population size of $10^4$
and recombination rate of $10^{-8}$) for varying (haploid) sample
size and sequence length. We ran 10 replicate mutation simulations
each for three different mutation rates, and report the average
CPU time required (Intel Core i7-9700).
(A) Holding sequence length fixed at 10 megabases and varying the
number of samples (tree tips) from 10 to 100,000.
(B) Holding number of samples fixed at 1000, and varying the sequence
length from 1 to 100 megabases.}
\end{figure}

Simulating mutations in \msprime\ is efficient.
Fig.~\ref{fig-mutations-perf} shows the time required to generate
mutations (using the default JC69 model) on
simulated tree sequences for a variety of mutation
rates as we vary the number of samples
(Fig.~\ref{fig-mutations-perf}A) and the sequence
length (Fig.~\ref{fig-mutations-perf}B).
For example, the longest running simulation in
Fig.~\ref{fig-mutations-perf}B required less than 2 seconds to
generate an average of 1.5 million mutations over 137,081 trees
in a tree sequence with 508,125 edges.
This efficiency for large numbers of trees is possible because
the tree sequence encoding allows us to generate mutations
on an edge-by-edge basis
(see Fig.~\ref{fig-ts-example} and the~\nameref{app-mutation-algorithm}
appendix),
rather than tree-by-tree and branch-by-branch as would otherwise be required.
In the above example from Fig.~\ref{fig-mutations-perf}B,
if we generated mutations tree-by-tree, we would have to iterate over 273,887,838 branches
(since there are 137,081 trees and 1,998 branches in each
tree) rather than 508,125 edges, resulting in $\sim$500 times more work.
% python3 evaluation/generate_mutations_perf_data.py benchmark-single-tree
Even if we have a tree sequence consisting of a single tree
(negating the advantage of working edge-by-edge),
\msprime's mutation generator is still very efficient.
For example, we simulated mutations under the BLOSUM62 amino
acid model for a tree with $10^6$ leaves over $10^4$ sites (resulting
in $\sim$260,000 mutations) in about $0.8$ seconds, including
the time required for file input and output.
We do not attempt a systematic benchmarking of \msprime's
mutation generation code against other methods, because at this scale it is
difficult to disentangle the effects of inefficient input and
output formats from the mutation generation algorithms.
Given these timings, it seems quite unlikely
that generating mutations with \msprime\ would be a bottleneck in any
realistic analysis.

There are many ways in which the mutation generation code
in \msprime\ could be extended. For example, we intend to add support for
microsatellites~\citep{mailund2005coasim},
codon models~\citep{arenas2007recodon}
and indels~\citep{cartwright2005dna,fletcher2009indelible},
although changes may be required to \tskit's data model
which is currently based on the assumption of independent sites.

\subsection*{Recombination}

Crossover recombination is implemented in \msprime\ using Hudson's algorithm, which
works backwards in time, generating common ancestor and recombination
events and tracking their effects on segments of ancestral
material inherited from the
sample~\citep{hudson1983properties,hudson1990gene,kelleher2016efficient}.
Common ancestor events merge the ancestral material of two lineages,
and result in coalescences in the marginal trees when ancestral
segments overlap.
Recombination events split the ancestral material for some lineage
at a given breakpoint, creating two independent
lineages. Using the appropriate data structures~\citep{kelleher2016efficient},
this process is much more efficient to simulate than either
methods based on the Ancestral Recombination Graph
formulation~\citep{griffiths1991two,griffiths1997ancestral}
or the left-to-right approach~\citep{wiuf1999recombination,wiuf1999ancestry}.
In \msprime\ 1.0 recombination rates can vary along a chromosome, allowing
us to simulate recombination hotspots and patterns
of recombination from empirical maps.
The implementation of recombination in \msprime\ is extensively validated
against analytical results~\citep{hudson1983properties,kaplan1985use}
and simulations by \ms, \msHOT\ and \SLiM.

The Sequentially Markovian Coalescent (SMC) is an approximation of the
coalescent with recombination~\citep{mcvean2005approximating,marjoram2006fast},
and was primarily motivated by the need to
simulate longer genomes than was possible using tools like \ms.
The SMC is a good approximation to the
coalescent with recombination when we have fewer than five sampled
genomes~\citep{hobolth2014markovian,wilton2015smc}, but the
effects of the approximation are less well understood for larger
sample sizes, and several approaches have been proposed
that allow simulations to more closely approximate the coalescent
with recombination~\citep{chen2009fast,wang2014new,staab2015scrm}.
The SMC and SMC' models are available
in \msprime\ 1.0. However, they are currently implemented using a
naive rejection sampling approach, and are somewhat slower
to simulate than the exact coalescent with recombination. These
models are therefore currently only appropriate for studying the
SMC approximations themselves, although we intend to
implement them more efficiently in future versions.

\begin{figure}
\begin{center}
    \includegraphics[width=\textwidth]{figures/ancestry-perf}
\end{center}
\caption{\label{fig-ancestry-perf}
    Running time for \msprime\
    for (A) small and (B) larger simulations on an Intel i7-6600U CPU.
    Each point is the run time of one simulation,
    for various values of effective population size ($N_e$),
    genome length in Morgans ($L$), and number of samples ($n$).
    Run time scales quadratically with the product of $N_e$ and $L$,
    shown on the horizontal axis. For example, (A) shows that
    1,000 samples of 1 Morgan-length chromosomes
    from a population of $N_e=2,000$ diploids
    would take about 2 seconds, and (equivalently) that the same number of
    0.01 Morgan segments with $N_e= 200,000$ would take the same time.
    Since recombination rate in these simulations was $10^{-8}$,
    $L$ is the number of base pairs divided by $10^8$.
    The black lines are quadratic fits separately in each panel
    and sample size. Vertical grey lines show the approximate values of
    $N_e L$ for chromosome 1 in three species, using values from
    the \stdpopsim\ catalogue~\citep{adrion2020community}.
}
\end{figure}

% Output from  python3 evaluation/plot.py ancestry-perf
% Predicted time for DroMel chr2L with n = 1000 = 107.85106995938845 hours
% Predicted time for DroMel chr2L with n = 100000 = -511.74273443828514 hours
% Predicted time for DroMel chr2L with n = 1000 = 177.40184421122058 hours
% Predicted time for DroMel chr2L with n = 100000 = 361.90415260982536 hours

In human-like parameter regimes and for large sample sizes,
\msprime's implementation of the
exact coalescent with recombination comprehensively outperforms
all other simulators, including those based on SMC
approximations~\citep{kelleher2016efficient}. However, it is important
to note that although the implementation of Hudson's algorithm is
very efficient, it is still quadratic in the population scaled recombination rate
$\rho = 4 N_e L$, where $L$ is the length of the genome in units of recombination distance.
This is because Hudson's algorithm tracks recombinations not only
in segments ancestral to the sample, but also between ancestral segments.
As mentioned above, common ancestor events in which the ancestral material
of two lineages is merged only result in coalescences in the marginal trees
if their ancestral segments overlap. If there is no overlap, the merged
segments represent an ancestral chromosome that is a genetic ancestor
of the two lineages, but not the most recent common genetic ancestor
at any location along the genome.
When this happens, the merged lineage carries ``trapped'' genetic material
that is not ancestral to any samples, but where recombinations can still occur.
The SMC approximations work by disallowing common ancestor events
that generate trapped material, greatly simplifying the process.
However, this also removes subtle long-range
correlations in the trees since there are many ways in which ancestry
segments can merge without overlapping.

For large $\rho$, recombination events in trapped ancestral material will
dominate, and so we can use this as a proxy for the overall number
of events in Hudson's algorithm.
\citet[Eq.~5.10]{hein2004gene} gave
\begin{equation}\label{eqn-hsw-bound}
\rho (\rho + 1) \left( \sum_{i=1}^{n-1} \frac{1}{i} \right)^2
\end{equation}
as an upper bound
on the number of recombination events within trapped ancestral material
in Hudson's algorithm for $n$ samples.
% Since the sum is well approximated by $\log{n}$,
% a natural conjecture is that the time complexity of Hudson's
% algorithm is $O(\rho^2 \log^2 n)$.
Fig.~\ref{fig-ancestry-perf} shows the observed run time for
simulations with a variety of population size, chromosome
length and sample sizes, and demonstrates that Eq.~\eqref{eqn-hsw-bound}
correctly predicts the quadratic dependence on $\rho$,
as previously conjectured~\citep[Fig.~2]{kelleher2016efficient}.
We also see that the dependence on $n$ is quite weak,
since increasing sample size 100-fold
only increases run time by a factor of 2 or so. However, the
$\log^2{n}$ factor implied by Eq.~\eqref{eqn-hsw-bound}
(the sum is a harmonic number, and approximated by $\log{n}$)
is not well supported by the data except possibly at very
large values of $\rho$, and so it seems a different dependence
on $n$ would be required to accurately predict simulation time for a
given $\rho$ and $n$.

Fig.~\ref{fig-ancestry-perf} is a useful yardstick, allowing us
to predict how long simulations should take for a wide range of
species. Taking a typical chromosome to be 1 Morgan in length,
these plots show, roughly, that simulating chromosome-length
samples from a population of thousands of individuals takes seconds,
while samples from a population of tens of thousands takes minutes.
Simulating whole chromosomes for many species is very fast,
with 1000 samples of chromosome 1 for
\emph{Arabidopsis thaliana} taking less than a second, and a few
minutes for dogs and humans. However, the dependence on $\rho$
\emph{is} quadratic, and if $\rho$ is sufficiently large simulations
may not be feasible. For example,
the \emph{Drosophila melanogaster} chromosome 2L
is about 23.5Mb long and an average recombination rate of around $2.4 \times 10^{-8}$,
so $L \approx 0.57$,
and with $N_e = 1.7 \times 10^6$ \citep{li2006inferring},
$N_e L \approx 10^6$, so extrapolating the curve in Fig.~\ref{fig-ancestry-perf}B
predicts that simulation would require around 177 hours.
For such large values of $\rho$  we recommend
users consider approximate simulations. Since \msprime\ does not currently
have efficient implementations of approximate coalescent with recombination
models, we recommend using SMC based methods such as \scrm\,
particularly for moderate sample sizes.
In practice, to predict the running time of a given
simulation in \msprime, we recommend that users
measure run time in a series of simulations with short genome lengths and
the desired sample size,
and then predict run time by fitting a quadratic curve to genome length
as in Fig.~\ref{fig-ancestry-perf}.
It is important to note that the quadratic curves in the two
panels of Fig.~\ref{fig-ancestry-perf} are different, and
predicting the run times of days-long simulations
using the timing of minutes-long runs is unlikely to be very accurate.

What about simulations with changing population size?
To understand how run time depends on demography
it helps to consider why run time is quadratic in $\rho$.
At any point in time, \msprime\ must keep track of some number of lineages,
each of which contains some number of chunks of genetic material.
Common ancestor events reduce the number of lineages,
and recombination events increase their number.
However, with long genomes
only a small fraction of the common ancestor events
will involve overlapping segments of ancestry,
and lead to coalescence in the marginal trees.
Such disjoint segments are often far apart (on average, about distance $L/2$),
and so recombine apart again immediately;
it is these large numbers of rapid and inconsequential events that
lead to the quadratic run time.
The maximum number of lineages occurs when
the increase and decrease in numbers of lineages due to
common ancestor and recombination events balance out;
to get an idea of run time we can estimate when this occurs.
Suppose that the maximum number of lineages is $M$;
at this time the rate of common ancestor events is $M (M-1) / (4 N_e)$
and the total rate of recombination is $M \ell$,
where $\ell$ is the mean length of genome carried by each lineage
(including ``trapped'' non-ancestral material).
At the maximum, coalescence and recombination rates are equal,
so a typical segment of ancestry will spend roughly half its time
in a lineage with at least one other such segment---and,
since such lineages carry at least two segments,
at most one-third of the lineages carry long trapped segments of ancestry.
Since the maximum number of lineages is reached
very quickly~\citep{nelson2020accounting},
this implies that $\ell \approx L / 6$.
Setting the rates of recombination and common ancestor events
to be equal and solving for $M$,
we find that $M$ is roughly equal to $L N_e$.
The number of lineages then decreases gradually from this maximum
on the coalescent time scale, and therefore over roughly $2 N_e$ generations.
Since the total rate of events when the maximum number of lineages is
present is roughly $L^2 N_e / 6$,
then the total number of events is proportional to $(L N_e)^2$---i.e.,
proportional to $\rho^2$.

What does this tell us about time-varying population sizes?
The argument above implies that the work is spread out relatively
evenly on the coalescent time scale.
Suppose that population size today is $N_1$,
while longer ago than $T$ generations ago it was $N_2$:
does the run time depend more on $4 N_1 L$ or $4 N_2 L$?
The answer depends on how $T$ compares to $N_1$: if $T/N_1$ is large,
then run time will be similar to a population of size $N_1$;
while if $T/N_1$ is small, it will be similar to a population of size $N_2$.
For instance, in many agricultural species $N_1 \propto 100$, while $N_2 \propto 10^5$,
and the run time will depend critically on $T$---in other words,
simulation will be quick in a species with a strong domestication bottleneck,
and slow otherwise.

\subsection*{Gene conversion}
\grgcomment{TODO: discuss Fig~\ref{fig-gc-perf} in text of GC section.}

%introduction
Gene conversion is a form of recombination that results in unidirectional transfer
of a short segment of genetic material,
for example between homologous chromosomes~\citep{chen2007gene}.
%description of ancestral process
Just as in crossover recombination, gene conversion events split how genetic material
is inherited along some lineage, creating two independent lineages.
In contrast to the previously described recombination process, where the ancestral
material to the left and right of a recombination breakpoint follows two
different ancestral lineages, gene conversion events distribute the genetic material
of a short segment to one ancestral lineage and the part to the left and right of this
segment to another lineage.
%Franz Baumdicker: Not sure if this next sentence is necessary,
% but it might be of interest for some people?
The resulting distribution of ancestral material after a gene conversion event
equals the distribution after two recombination events at the left and right position
of the gene conversion tract, followed by a common ancestor event.
%Impact on LD
However, gene conversions impact much shorter segments than crossover recombinations,
typically below 1kb, and thus affect linkage patterns differently to
recombinations~\citep{korunes2017gene}.

% other simulators
Hudson's \ms\ optionally includes gene conversion and so does
MaCS~\citep{chen2009fast}, while scrm~\citep{staab2015scrm} does not
support gene conversions.
As bacterial recombination can be modelled analogously to eukaryotic allelic
gene conversion, simulators for microbial evolution often include gene conversion
instead of recombination.
This includes the backward simulator SimBac~\citep{brown2016simbac},
the SMC based approach used in FastSimBac~\citep{demaio2017the},
and forward-in-time tools within the SLiM framework~\citep{cury2020simulation}.

In \msprime\ 1.0 we implemented the coalescent with gene conversion as defined by
\cite{wiuf2000coalescent}, where gene conversion events are initiated between any
two sites in the sequence at a constant rate $\gamma$ and
the length of the transferred segment is given by a geometric distributed
random variable with mean length $L$.
% Above I used the discrete case, but we could easily switch to a description for the
% continuous case.
We have implemented this model of gene
conversion in \msprime\ 1.0, and validated the output against
\ms\ and analytical results~\citep{wiuf2000coalescent}.

% In terms of file size, as expected from the discussion above, we observe that gene
% conversion is equivalent to double the rate of
% crossover recombination, so that a tree sequence produced with a `recombination_rate` of $r$
% and a `gene_conversion_rate` of $g$ is roughly the same size as a tree sequence with no gene
% conversion and a `recombination_rate` of $r+2g$. However, because gene conversion tends to
% affect only short regions of the genome, each gene conversion event usually creates
% less trapped material, hence simulations using mammalian-like gene conversion parameters
% run faster than simulations in which an equivalent amount of crossover recombination is
% imposed.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/gc-perf}
\end{center}
\caption{\label{fig-gc-perf}Comparison of the running time
for E-coli simulations using \msprime\ and SimBac for varying
sample sizes. The parameters used are chosen to mimic the simulation
of \textit{E. coli} genomes. The genome length was set to 4.5 Mb, the scaled per site
gene conversion rate was set to $N_e g = \gamma = 0.015$ and
the mean gene conversion tract length was set to $L = 500$, which is close to
estimated parameters for \textit{E.\ coli}, i.e.\ an effective populations size
$N_e = 1.8e^8$ and a per site per generation conversion rate $g = 8.9e^{-11}$
with mean tract length $L = 542$~\citep{lapierre2016the}.}
\end{figure}

Gene conversion is particularly useful to model homologous recombination in
bacterial evolution,
so we compare the performance of \msprime\ with gene conversion to
SimBac~\citep{brown2016simbac}, a simulator developed specifically to
simulate bacterial evolution. Although \msprime\ is missing some features
relevant to this case, e.g.\ the simulation of circular genomes, it is far
more efficient at simulating the basic processes.
Furthermore, gene conversion can be considered in combination
with other current and future features of \msprime\ 1.0,
including varying recombination rates and complex demographies.

An important avenue for future work
will be to add further features to model bacterial
evolution. In particular, we aim to include bacterial gene transfer and utilise
the succinct tree sequence structure to represent the resulting ancestral gene
transfer graph~\citep{baumdicker2014AGTG}.

\subsection*{Demography}
% TODO should we put in some citations here for the models? Hardly seems
% necessary.
One of the key applications of population genetic simulations is to generate
data for complex demographies. Beyond idealised cases such as stepping-stone or
island models, or specialised cases such as isolation-with-migration models,
analytical results are rarely possible. Simulation is therefore integral to the
development and evaluation of methods for demographic inference. The demography
model in \msprime\ is directly derived from the approach used in \ms, and
supports an arbitrary number of randomly mating populations exchanging migrants
at specified rates. A range of demographic events are supported, which allow for
varying population sizes and growth rates, changing migration rates over time,
as well as population splits, admixtures and pulse migrations.
The location of sampled lineages
can be tracked through time in as much detail as required: each tree node is
automatically associated with the population in which it arose, the location of
lineages can be recorded at any given time via census events, or every lineage
migration can be recorded.

A major change for \msprime\ 1.0 is the introduction of the new Demography API,
designed to address a design flaw in the \msprime\ 0.x interface which lead to
a number of avoidable errors in downstream
simulations~\citep{ragsdale2020lessons}. Briefly, the 0.x API required three
separate parameters be provided to the \texttt{simulate} function to describe a
demographic model, making it easy to accidentally omit information. The 1.0 API
resolves this issue by creating a new \texttt{Demography} class, which
encapsulates all information about the demographic model, and fully decouples
the definition from other simulation details. An instance of this class is then
provided as a parameter to the new \texttt{sim\_ancestry} function,
substantially reducing the potential for error.
Another improvement over the
0.X APIs is the introduction of explicit population split and admixture events,
and a population state machine that ensures that lineages cannot migrate into
(or be sampled from) inactive populations. This demography model is compatible
with the Demes standard~\citep{gower2021demes}, and the \texttt{Demography}
class supports importing and exporting Demes models.
Models previously constructed using the 0.x API can be seamlessly imported into
the \texttt{Demography} class, and we also support importing demographic
models in the form of Newick species tree descriptions, as well as those
produced by programs such as *BEAST~\citep{heled2009bayesian}.

The \texttt{DemographyDebugger} can output detailed information for debugging
demographic models, but also provides
methods to compute a suite of numerical predictions about the location of
lineages over time, for example the coalescence rates for two or more lineages
drawn from populations at specified times in the past,
which can be inverted to obtain the ``inverse instantaneous coalescence rate'',
or IICR of \citet{chikhi2018iicr}.
Many popular approaches in population genetics use the distribution of
coalescence rates between pairs of lineages in one or more populations to infer
effective population sizes over
time~\citep{li2011inference,sheehan2013estimating,schiffels2014inferring} or
split times and subsequent migration rates between
populations~\citep{wang2020tracking}.
The \texttt{DemographyDebugger} provides a valuable
ground-truth when evaluating such inference methods, as illustrated by
\cite{adrion2020community}.

\subsection*{Instantaneous bottlenecks}

A common approach to modelling the effect of demographic history on genealogies
is to assume that effective population size ($N_e$) changes in discrete steps
which define a series of epochs~\citep{griffiths1994sampling, marth2004allele,
keightley2007joint,li2011inference}. In this setting of piecewise constant
$N_e$, capturing a population bottleneck requires three epochs: $N_e$ is
reduced by some fraction $b$ at the start of the bottleneck, $T_{start}$, and
recovers to its initial value at time $T_{end}$~\citep{marth2004allele}. If
bottlenecks are short both on the timescale of coalescence and mutations,
there may be little information about the duration of a bottleneck
($T_{end}-T_{start}$) in sequence data. Thus a simpler, alternative model is to
assume that bottlenecks are instantaneous ($T_{end}-T_{start} \rightarrow 0$)
and generate a sudden burst of coalescence events (a multiple merger event) in
the genealogy. The strength of the bottleneck $B$ can be thought of as an
(imaginary) time period during which coalescence events are collapsed,
i.e.\ there is no growth in genealogical branches during $B$ and the probability that a
single pair of lineages entering the bottleneck coalesce during the bottleneck
is $1-e^{-B}$. Although this simple two parameter model of bottlenecks is
attractive and both analytic results and empirical
inference~\citep{griffiths1994sampling, galtier2000detecting,
bunnefeld2015inferring} have been developed under this model, there has
been no software available to simulate data under instantaneous
bottleneck histories.

We have implemented instantaneous bottlenecks in \msprime~1.0
using a variant of Hudson's linear time single-locus coalescent
algorithm~\citep{hudson1990gene}. Instantaneous bottlenecks are specified
by adding events to the Demography object (see the Demography section)
and can be used in combination with any other demographic modelling
features. We have validated the results of these simulations by comparing
against analytical expectations for coalescence times and the
site frequency spectrum~\citep{bunnefeld2015inferring}.

\subsection*{Multiple merger coalescents}

Kingman's coalescent assumes that only two ancestral lineages can merge at
each merger event. Although this is generally a reasonable approximation, there
are certain situations in which the underlying mathematical assumptions are
violated. For example in certain highly fecund organisms
\citep{hedgecock_94,B94,HP11,A04,irwin16}, where individuals have the ability
to produce numbers of offspring on the order of the population size and
therefore a few individuals may produce the bulk of the offspring in any given
generation~\citep{hedgecock_94}. These population dynamics violate basic
assumptions of the Kingman coalescent, and are better modelled by
`multiple-merger' coalescents \citep{DK99,P99,S99,S00,MS01}, in which more than
two lineages can merge in a given event. Multiple-merger coalescent processes
have also been shown to be relevant for modelling the effects of selection on
gene genealogies \citep{Gillespie909,DS04}.

Although multiple merger coalescents have been of significant theoretical
interest for around two decades, there has been little practical software
available to simulate these models.
\cite{kelleher2013coalescent,kelleher2014coalescent} developed packages to
simulate a related spatial continuum model~\citep{barton2010new},
\cite{zhu2015hybrid} simulate genealogies within a species tree
based on a multiple-merger model, and
\cite{becheler2020occupancy} provide a general method for simulating
multiple merger processes
as part of the Quetzal framework~\citep{becheler2019quetzal}.
The \texttt{Beta-Xi-Sim} simulator~\citep{koskela2018multi,koskela2019robust}
includes a number of extensions to the Beta-coalescent.
None of these methods work with large genomes, and very little work
has been performed on simulating multiple merger processes with recombination.

We have added two multiple merger coalescent models in \msprime\ 1.0, the
Beta-coalescent and ``Dirac''-coalescent, allowing us to simulate
such models with recombination efficiently for the first time.
These simulation models have been extensively validated against
analytical results from the site frequency
spectrum~\citep{birkner2013statistical,blath2016site,hobolth2019phase}
 as well as more general properties of coalescent processes.
Please see the Appendix for more details and model derivations.

\subsection*{Ancestral Recombination Graphs}
\label{sec-arg}

The Ancestral Recombination Graph (ARG) was introduced by
Griffiths~\citep{griffiths1991two,griffiths1997ancestral}
as a representation of the coalescent with recombination in terms of a graph.
This formulation was complementary to Hudson's earlier
work~\citep{hudson1983properties}, and substantially increased our theoretical
understanding of recombination. In Griffiths' ARG formulation,
a realisation of the coalescent with recombination is a graph in which
vertices represent common ancestor or recombination events, and edges represent
lineages through which ancestral material flows.
We may consider the big ARG, in which we track lineages arising out of
recombinations regardless of whether they carry ancestral material and
consequently wait an exponentially large time until all sampled lineages
coalesce into a single ancestor~\citep{ethier1990two}, or the little ARG in
which we only track genetic ancestors, and wait until all marginal trees
have fully coalesced (like Hudson's algorithm).

\begin{figure}

\begin{center}
    \begin{tikzpicture}[scale=0.45]

    \draw (-2,0) -- (-2,3) -- (0, 3);
    \draw (1,0) -- (1,1);
    \draw (0,1) -- (2,1) -- (2,2) -- (4,2) -- (4,0);
    \draw (3,2) -- (3,5);
    \draw (0,1) -- (0,3);
    \draw (-1, 3) -- (-1, 5) -- (3,5);

    \draw[fill] (-2,0) circle [radius=0.05];
    \draw[fill] (-1,3) circle [radius=0.05];
    \draw[fill] (1,0) circle [radius=0.05];
    \draw[fill] (4,0) circle [radius=0.05];
    \draw[fill] (1,1) circle [radius=0.05];
    \draw[fill] (3,2) circle [radius=0.05];
    \draw[fill] (1,5) circle [radius=0.05];

    \node [below] at (-2,0) {0};
    \node [below] at (1,0) {1};
    \node [below] at (4,0) {2};
    \node [above] at (1,1) {3};
    \node [below right] at (1,1) {$x = 0.3$};
    \node [above right] at (3,2) {4};
    \node [above left] at (-1,3) {5};
    \node [above] at (1,5) {6};

    \end{tikzpicture}
% \begin{center}
% $\equiv$
% \end{center}
% \end{column}
% \begin{column}{5cm}
\includegraphics[width=3.8cm]{illustrations/arg-ts.pdf}
\includegraphics[width=3.8cm]{illustrations/arg-ts-simplified.pdf}
\includegraphics[width=3.8cm]{figures/arg.pdf}

\end{center}
\caption{\label{fig-arg} (A) A simple ARG in which a recombination
occurs at position 0.3; (B) the equivalent topology depicted as a tree
sequence, including the recombination node; (C) the same tree sequence
topology ``simplified'' down to its minimal tree sequence representation;
(D) the fraction of nodes in a tree sequence that are concerned with
extra ARG information in a tree sequence returned using \msprime's
\texttt{record\_full\_arg} mode}
\end{figure}

However, this original definition of the ARG in terms of a stochastic process
is rarely what is meant when the term is currently used. The term has come to
essentially mean any graph-like representation of genetic ancestry in the presence
of recombination; for example,
\cite{mathieson2020ancestry} informally define an ARG as the subset of an
individual's pedigree from which they have inherited ancestral material.
Others state that an ARG is equivalent to the sequence of marginal
% Also SARGE paper
trees along the genome~\citep{yang2016rapidly,mirzaei2016rent}

Such
imprecision is unfortunate, since there are many ways to encode genetic
ancestry as a graph, and referring to the original Griffiths papers does little to
clarify. As we see below, there are situations in which the precise information
stored in the graph is important, and conflating a high-level
discussion of genetic ancestry with this mathematical object is not
helpful. The point of technical terms is to provide clarity and
precision on what is being discussed, and it is now essentially
impossible to know what is meant by the term ``ARG'';
we therefore use the term ``genetic ancestry'' in high-level
discussions.

For our purposes, an ARG is a realisation of the coalescent with
recombination, in the Griffiths (little ARG) sense.
As described in detail by~\cite{kelleher2016efficient}, Hudson's algorithm
works by dynamically generating and traversing a little ARG.
The graph is not explicitly represented in memory, but is partially
present through the extant lineages and the ancestral material they carry
over time. Since the ARG is a large and unwieldy object
(Fig.~\ref{fig-arg}) we do not use it as output, but
rather store the information required to recover the genealogical
history as nodes and edges in a tree sequence (see the \nameref{sec-ts} section),
or ``coalescence records'' in the terminology of~\citep{kelleher2016efficient}.

Although by default \msprime\ outputs tree sequences that contain full information
about the genealogical trees, their correlation structure along the genome,
and the ancestral haplotypes on which coalescences occurred, some information
is lost in this mapping down from ARG space to the minimal tree sequence
form. In particular, we lose
information about ancestral haplotypes that were common ancestors but
in which no coalescences occurred, and also information about the precise time
and genomic location of recombination events. In most cases, such
information is of little relevance as it is in principle unknowable,
but there are occasions such as visualisation or computing likelihoods (see
below) in which is useful. We therefore provide
the \texttt{record\_full\_arg} option in \msprime\
to store a representation of the complete ARG traversed during simulation.
This is done by storing extra nodes (marked with specific flags, so they
can be easily identified later) and edges in the tree sequence.
Fig.~\ref{fig-arg} shows that this extra ARG information
quickly comes to dominate the output as we increase the sequence length,
and the resulting tree sequence files are many times larger
than the standard minimal tree sequence representation.

One situation in which a record of the full ARG is necessary is when we
wish to compute likelihoods during inference.
The likelihood is a central quantity in evaluating the plausibility of a putative
ancestry as an explanation of DNA sequence data, both directly through
e.g.~approaches based on maximum likelihood, and as an ingredient of
methods such as
Metropolis-Hastings~\citep{kuhner2000maximum,nielsen2000estimation,wang2008bayesian}.
We provide functions to compute the likelihood of ARG realisations
(encoded using the representation of the \texttt{record\_full\_arg} option)
and mutational patterns under
the standard coalescent and infinite sites mutation model.
See the Appendix for details.

\subsection*{Selection}
\label{sec-selection}
Another elaboration of the standard neutral coalescent with recombination
is the addition of selective
sweeps~\citep{kaplan1989hitchhiking,braverman1995hitchhiking,kim2002detecting}.
Sweeps are modelled by creating a structured population
during the sojourn of the beneficial mutation through the population (i.e., the sweep phase)
in which lineages may transit between favoured and unfavoured backgrounds through
recombination. This allows for many selective sweep scenarios to be simulated
efficiently, including recurrent, partial, and soft selective sweeps.
However in general this
efficiency comes at the cost of flexibility in comparison to forwards in
time simulation.
Several specialised simulators have been developed to simulate
sweeps in the coalescent,
including \texttt{SelSim}~\citep{spencer2004selsim},
\texttt{mbs}~\citep{teshima2009mbs}, \msms~\citep{ewing2010msms},
\cosiTwo~\citep{shlyakhter2014cosi2}
and \discoal~\citep{kern2016discoal}.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/sweeps-perf}
\end{center}
\caption{\label{fig-selection-perf}Comparison of selective sweep simulation
performance in \msprime\ and \discoal\ (Intel Xeon E5-2680 CPU).
We report the total CPU time and maximum memory usage when simulating 100
replicates for 10 samples in a model
with \jkcomment{[ADK: fill in the details here please]} and a
per-base recombination rate of $10^{-9}$.}
\end{figure}

Following the approach taken in \discoal, we
have implemented selective sweeps in \msprime\ by generating
sweep trajectories from a jump process approximation to the conditional diffusion
of an allele bound for
fixation~\citep{coop2004ancestral}---see the appendix for more details.
This implementation facilitates future changes to generalise sweep models to
changing population sizes as well as selection from standing variation,
both of which are features that are planned for the near future.
Moreover because this implements the basic infrastructure for the structured
coalescent, adding support for similar situations,
such as inversions~\citep{peischl2013sequential}, is straightforward.
Fig.~\ref{fig-selection-perf} shows that \msprime\ has far better
CPU time and memory performance than \discoal. We have previously shown
that \msprime\ is much more efficient than \msms\ and \cosiTwo\
when simulating without selection~\citep{kelleher2016efficient}.
% JK: do we need an extra proviso like this?
% and this is unlikely to change unless sweeps dominate the ancestral process.

\subsection*{Discrete time Wright-Fisher}
The coalescent is an idealised model and makes many simplifying assumptions,
but it is often surprisingly robust to violations of these
assumptions~\citep{wakeley2012gene}. One situation in which the
model does break down is the combination of large sample size
and with long recombining genomes, where the large
number of recombination events in the recent past results in
more than the biologically possible $2^t$ ancestors in
$t$ diploid generations~\citep{nelson2020accounting}.
This pathological behaviour results in
Identity-by-descent,
long-range linkage disequilibrium (LD), and ancestry patterns deviating from
Wright-Fisher expectations, and the bias grows with larger sample
sizes~\citep{wakeley2012gene,bhaskar2014distortion,nelson2020accounting}.
Precisely this problem occurs when simulating modern human datasets,
and we have implemented a Discrete Time Wright-Fisher (DTWF) model
in \msprime\ to address the issue. The DTWF simulates backwards in
time generation-by-generation so that each gamete has a unique
diploid parent, and multiple recombinations within a generation results in
crossover events between the same two parental haploid copies.
The method is described in more detail by~\cite{nelson2020accounting}.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/dtwf-perf}
\end{center}
\caption{\label{fig-dtwf-perf} Comparison of Discrete Time Wright-Fisher
(DTWF) simulation performance in \msprime\ and \ARGON\ (Intel Xeon E5-2680 CPU).
We simulated ancestry for a sample of 1000 haploids from a population of 10000,
and report the (A) total CPU time and (B) maximum memory usage for varying
sequence lengths, and a per-base recombination rate of $10^{-8}$. Each
point is the average over $5$ replicate simulations.
We show observations for \ARGON, \msprime's DTWF implementation (``DWTF'')
and a hybrid simulation of 100 generations of the DTWF followed by
the standard coalescent with recombination model (``DTWF + Hudson'').
We ran \ARGON\ with a mutation rate of $0$ and with minimum output options,
to ensure we are measuring only ancestry simulation time.}
\end{figure}

Fig.~\ref{fig-dtwf-perf} shows that \msprime\ simulates the DTWF
more quickly and requires substantially less memory than
\ARGON~\citep{palamara2016argon}, a specialised DTWF simulator.
However, the generation-by-generation approach of the DTWF is less
efficient than the coalescent with recombination when the
number of lineages is significantly less than the population size
(the regime where the coalescent is an accurate approximation),
which usually happens in the quite recent
past~\cite{bhaskar2014distortion}.
We therefore support changing the simulation model during a simulation
so that we can run hybrid simulations, as proposed by~\cite{bhaskar2014distortion}.
Any number of different simulation models can be combined, allowing for the
flexible choice of simulation scenarios.
As the discrete time Wright-Fisher model improves accuracy of genealogical
patterns in the recent past, we can simulate the recent history using this
model and then switch to the standard coalescent to more efficiently simulate
the more ancient history.

\subsection*{Integration with forward simulators}
A unique feature of \msprime\ is its ability to simulate
genetic ancestries by extending an existing partial
genetic ancestry. Given a tree sequence that
is complete up until time $t$ as input
(where marginal trees may or may not have fully coalesced),
\msprime\ can efficiently obtain the segments of ancestral material
present at this time, and then run the simulation backwards in time from there.
This allows a simulated ancestry to be produced by any
number of different processes across disjoint time slices.
In practice this feature is used to ``complete''
forwards-time ancestry simulations~\citep{kelleher2018efficient}
that may have not fully coalesced. This process
(``recapitation'') can be orders of magnitude faster than
the standard approach of neutral burn-in; see
\cite{haller2018tree} for more details and examples.
This interoperability between simulators, where a partial ancestry
simulation produced by \SLiM~\citep{haller2019slim}
or \fwdpy~\citep{thornton2014cpp} can be picked up and completed
by another simulator, with complete information
retained---at scale---is unprecedented.

\subsection*{Development model}
\texttt{Msprime} has a large number of features, encompassing
the functionality of several more specialised simulators
while maintaining excellent performance.
It is developed by a geographically distributed team of volunteers under an
open source community development model, with a strong emphasis
on code quality, correctness, good documentation, and inclusive development.
As in any large code base,
unit tests play a key role in ensuring that new additions behave
as expected and \msprime\ has an extensive suite.
As of the 1.0.0 release \msprime\ consists of around 13K lines of C and 11K
lines of Python, with suites of 122 C tests (7K lines of code) and
1350 Python tests (22K lines of code). These tests are run
automatically on different operating systems on each pull request
(where a contributor proposes a code change), using standard Continuous
Integration (CI) methodology. Other CI services check code formatting
and for common errors, as well as producing reports on the level of
test coverage, e.g., pinpointing areas where additional unit tests may
be required.

Unit tests are vital for ensuring software quality and correctness, but
they are usually of little value in assessing the statistical properties
of simulations. To validate the correctness of simulation output
we maintain a suite of statistical tests (as of 1.0.0,
217 validation tests, in 6K lines of code). These consist of running many
replicate simulations to check the properties of the output
against other simulators, and where possible against analytical results.
For example, simulations of complex demography are validated against \ms,
selective sweeps against \discoal, and Wright-Fisher simulations
against forwards in time simulations in \SLiM.
This suite of tests is run before every release, to ensure
that statistical errors have not been introduced.

More visibly to the end user, we also have a high standard for documentation,
with precise, comprehensive, and cross-linked documentation
that is automatically built from the code base
and served through the website \texttt{https://tskit.dev}.
With the goal of lowering the entry barrier to new users,
we have invested significant effort in writing examples and introductions,
and making common tasks discoverable.
We also view contributions to documentation as equally important to the project
as writing code or designing methods---what use would it be
to write reliable, stable software if no-one used it?

\section*{Discussion}

% We can simulate so much more now

This release makes \msprime\ able to produce simulations
with a great deal more biological realism, without a loss in speed and efficiency.
This ability will allow researchers to perform more robust
power analyses, more reliably test new methods,
and carry out more reliable inferences,
without worrying that gene conversion or back mutation might be sabotaging
their applications to real data.
These advances have only been possible thanks to a distributed,
collaborative model of software development,
and the work of many people.
We hope the tool continues to be useful to many.

% Simulation software is historically a beautiful shitshow

Even though simulation has long been a vital tool in population genetics,
such collaborative software development has historically been uncommon.
A huge proliferation of tools have been published
(the references here are not exhaustive)
and only a small minority of these are actively developed and
maintained today. The ecosystem is highly fragmented, with numerous different
ways of specifying parameters and representing results, and there are
significant software quality issues at all stages. This is unsurprising, since
the majority of simulation software development is performed by students, who
are themselves learning to code, often without formal training in software
development.
Although this ``Haldane's sieve'' of software has produced many good tools
and enabled decades of research,
it also represents a missed opportunity to invest as a community
in shared infrastructure and mentorship in good software development practice.

% THERE IS A BETTER WAY!!!
Scientific software is vital apparatus, and must be engineered
to a high quality if we are to trust its results. There is a growing
realisation across the sciences \citet[OTHER CITATIONS]{siepel2019challenges}
% Maybe cite: https://www.biorxiv.org/content/10.1101/092205v3
that investing in
shared community infrastructure produces better results than a
proliferation of individually maintained tools, allowing scientists
to focus on their specific questions rather than software engineering.
\texttt{Msprime} 1.0 is the result of such a community process,
with features added by motivated users, taking advantage of the
established development practices and infrastructure.
% I'm going out on a limb here - do folks agree?
% PLR- personally I have learned SO MUCH, and I'm passing it on to others
Software development in a welcoming community,
with mentorship by experienced developers,
is a hugely useful experience for many users,
The skills that contributors learn
can lead to greatly increased productivity in subsequent
work (e.g., through more reliable code and better debugging skills).
We hope that users who find that
features they require are missing will continue to contribute to
\msprime, leading to a community project that is both high quality
and sustainable in the long term.

The tree sequence developed for \msprime\ allows users a view of not only
genetic variation but also the genetic ancestry that produced that variation.
Recent breakthroughs in methods to infer genetic ancestry in recombining
organisms~\citep{kelleher2019inferring,
speidel2019method,wohns2021unified,schaefer2021ancestral,speidel2021inferring}
have made it possible to estimate such ancestry from real data at scale for
the first time~\citep{harris2019database,tang2019genealogy}.
Given such inferred ancestry, many exciting applications
become possible. For example, \cite{osmond2021estimating}
developed a method to estimate the location of genetic ancestors
based on inferred trees, and other uses are sure to follow.
Since the inferred genetic ancestry
becomes the input for other downstream inferences, it is vitally
important that these primary inferences are thoroughly validated,
with the detailed properties of the inferred ancestries
catalogued and understood.
\texttt{Msprime} will continue to be an important tool for these inferences and validations,
and in this context the ability to interoperate
with other methods---particularly forwards simulators---through
the succinct tree sequence data structure and \tskit\ library
will be essential.
% this final sentence is hard to parse

\section*{Acknowledgements}
JK is supported by the Robertson Foundation.
Jere Koskela is supported in part by EPSRC grant EP/R044732/1.
ADK was supported by NIH awards R01GM117241 and R01HG010774.
Franz Baumdicker is funded by the Deutsche Forschungsgemeinschaft
EXC 2064/1 -- Project number 390727645, and EXC 2124 -- Project number 390838134.
Graham Gower was supported by a Villum Fonden Young Investigator award to Fernando Racimo (project no. 00025300).

\bibliographystyle{plainnat}
\bibliography{paper}

%% local definitions for section multiple merger coalescents
 \newcommand{\be}{\begin{equation}}
 \newcommand{\ee}{\end{equation}}
 \newcommand{\bd}{\begin{displaymath}}
 \newcommand{\ed}{\end{displaymath}}
\newcommand{\IN}{\ensuremath{\mathds{N}}}%
\newcommand{\EE}[1]{\ensuremath{\mathds{E}\left[ #1 \right]}}%
\newcommand{\one}[1]{\ensuremath{\mathds{1}_{\left\{ #1 \right\}}}}%
\newcommand{\prb}[1]{\ensuremath{\mathds{P}\left( #1 \right) } }%

\NewEnviron{esplit}[1]{%
\begin{equation}
\label{#1}
\begin{split}
  \BODY
\end{split}\end{equation}
}

\setcounter{secnumdepth}{2} % Print out appendix section numbers

\appendix

\section*{Mutation generation}
\label{app-mutation-algorithm}

The algorithm that \msprime\ uses to simulate mutations on a tree sequence
proceeds in two steps:
first, mutations are ``placed'' on the tree sequence
(i.e., sampling their locations in time, along the genome, and on the marginal tree),
and then the ancestral and derived alleles of each mutation are generated.
All mutation models share the code to place mutations,
but choose alleles in different ways.

First, mutations are placed on the tree sequence under an inhomogeneous Poisson model
by applying them independently to each edge.
If an edge spans a region $[a, b)$ of the genome
and connected parent and child nodes with times $s < t$,
and the mutation rate locally is $\mu$,
then the number of mutations on the edge is Poisson with mean $\mu (t-s) (b-a)$,
and each mutation is placed independently at a position chosen uniformly in $[a,b)$
and a time uniformly in $[s,t)$.
In a discrete genome, all positions are integers and so more than one mutation
may occur at the same position on the same edge.
Otherwise (i.e., for an infinite-sites model),
positions are rejection sampled to obtain a unique floating-point number.
If an edge spans a region of the genome with more than one mutation rate,
this is done separately for each distinct region.
The algorithm scales with the number of edges in the tree sequence.
Since the number of edges is proportional to the number of marginal trees,
an algorithm that applies mutations separately to each marginal tree
would take at least as many steps,
and would require additional bookkeeping necessary to place mutations uniformly on a tree.

Next, alleles are chosen for each mutation.
If the site was not previously mutated, then a new ancestral allele is chosen for the site,
according to an input distribution of ancestral state allele probabilities.
Then, each mutation on the tree is considered in turn,
in order so that older mutations are considered first,
and a derived allele is chosen based on the parental allele
(which may be the ancestral allele or the derived allele of a previous mutation)
and some randomness.
Finally, information about the mutations are recorded in the site and mutation tables
of the tree sequence.

A mutation model must, therefore, provide two things:
a way of choosing an ancestral allele for each new variant site,
and a way of choosing a derived allele given the parental allele at each mutation.
Perhaps the simplest mutation model implemented in \msprime\ is the
\texttt{InfiniteAlleles} mutation model,
which keeps an internal counter so that the requested alleles are
assigned subsequent (and therefore unique) integers.

Since the algorithm above applies mutations at a single rate
independent of ancestral state,
a model in which different alleles mutate at different rates
must necessarily produce some silent mutations,
i.e., mutations in which the derived allele is equal to the parental allele.
To illustrate this, consider a mutation model on the binary alphabet
$\{X, Y\}$ in which $X$ mutates to $Y$ at rate $\alpha$
and $Y$ mutates to $X$ at rate $\beta$, with $\beta < \alpha$.
To implement this, first place mutations at the largest total rate, which is $\alpha$.
Then, at each site, choose an ancestral allele from the ``root distribution''
(a natural choice would be to choose from the stationary distribution,
$X$ with probability $\beta / (\alpha + \beta)$),
and for each mutation, choose a derived allele as follows:
if the parental allele is $X$, then choose $Y$;
if the parental allele is $Y$, then choose $X$ with probability $\beta/(\alpha + \beta)$,
and choose $Y$ otherwise.
This produces the correct distribution by Poisson thinning:
a Poisson process with rate $\alpha$ in which each point is discarded independently
with probability $\beta / (\alpha + \beta)$ is equivalent to a Poisson
process with rate $\beta$.
All finite-state models (implemented under the generic \texttt{MatrixMutationModel} class)
work in this way: mutations are placed at the maximum mutation rate,
and then some silent mutations will result.

In previous versions of \msprime, silent mutations were disallowed,
and we could have removed them from the output entirely.
However, we have chosen to leave them in, so that for instance
simulating with the \texttt{HKY} mutation model will result in silent mutations
if not all equilibrium frequencies are the same.
Although the presence of silent mutations may at first be surprising
(although they don't affect calculations of statistics in any way),
there is a good reason to leave them in:
to allow layering of different mutation models.
Suppose that we wanted to model the mutation process as a mixture of more than one model,
e.g., Jukes-Cantor mutations at rate $\mu_1$, and
HKY mutations occur at rate $\mu_2$.
Layering multiple calls to \texttt{sim\_mutations} is allowed,
so we could first apply mutations with the \texttt{JC69} model at rate $\mu_1$
and then add more with the \texttt{HKY} model at rate $\mu_2$.
However, there is a small statistical problem:
suppose that after applying Jukes-Cantor mutations we have an $A \to C$ mutation,
but then the HKY mutations inserts another mutation in the middle,
resulting in $A \to C \to C$.
If neither mutation model allows silent transitions,
then this is clearly not correct,
i.e., it is not equivalent to a model that simultaneously applies the two models.
(The impact is small, however, as it only affects sites with more than one mutation.)
The solution is to make the Jukes-Cantor model \emph{state-independent}
(also called ``parent-independent''),
by placing mutations at rate $4/3 \mu_1$ and then choosing the derived state for each mutation
\emph{independently} of the parent (so that 1/4 of mutations will be silent).
If so -- and, more generally, if the first mutational process put down is state-independent --
then the result of sequentially applying the two mutation models
is equivalent to the simultaneous model.
To facilitate this, many mutation models have a \texttt{state\_independent} option
that increases the number of silent mutations
and makes the model closer to state-independent.

\section*{Likelihood calculations}
\label{app-likelihoods}

We provide two functions to facilitate likelihood-based inference.
Both are implemented only for the simplest case of the standard ARG with a
constant population size, and require tree sequences compatible with the
\texttt{record\_full\_arg} option as their arguments.
\begin{itemize}
\item \texttt{msprime.log\_arg\_likelihood(ts, r, N)} returns the natural logarithm of
the sampling probability of the tree sequence \texttt{ts} under the ARG with per-link,
per-generation recombination probability \texttt{r} and population size \texttt{N}.
Specifically, the function returns the logarithm of
\begin{equation*}
\Bigg( \frac{ 1 }{ 2 N } \Bigg)^{ q_c } \Bigg( \prod_{ i : \mathcal{R} } r g_i \Bigg)
	\exp\Bigg( -\sum_{ i = 1 }^q \Big[\frac{ 1 }{ 2 N } \binom{ k_i }{ 2 }
		+ r l_i \Big] t_i  \Bigg),
\end{equation*}
where $t_i$ is the number of generations between the $(i - 1)$th and $i$th event,
$k_i$ is the number of extant ancestors in that interval, $l_i$ is the number of links
in that interval that would split ancestral material should they recombine,
$q$ is the total number of events in the tree sequence $\texttt{ts}$,
$q_c$ is the number of coalescences, $\mathcal{R}$ is the set of indices
of time intervals which end in a recombination, and $g_i$ is the corresponding
\emph{gap}:  the length of material in which the recombination at the end of the
$i$th period could have taken place without affecting the resulting genealogy.
A recombination in ancestral material will always have $g_i = 1$, while a recombination
in trapped non-ancestral material can result in other values of $g_i$.
For a continuous model of the genome and a recombination in ancestral material,
we set $g_i = 1$ and interpret the result as a density.
\item \texttt{msprime.unnormalised\_log\_mutation\_likelihood(ts, m)} returns the
natural logarithm of the probability of the mutations recorded in the tree sequence
\texttt{ts} given the corresponding ancestry, assuming the infinite sites model, up to
a normalising constant which depends on the pattern of mutations,
but not on the tree sequence or the per-site, per-generation mutation
probability \texttt{m}.
Specifically, the function returns the logarithm of
\begin{equation*}
e^{ - T m / 2 } \frac{ ( T m / 2 )^M }{ M ! }
\prod_{ \gamma \in \mathcal{ M } } \frac{ h_{ \gamma } }{ T },
\end{equation*}
where $T$ and $\mathcal{M}$ are the total branch length and set of mutations
in \texttt{ts}, respectively, and for a mutation $\gamma$, $h_{ \gamma }$ is the
total branch length on which $\gamma$ could have arisen while appearing on all
of the leaves of \texttt{ts} it does, and on no others.
Unary nodes on marginal trees arising from the \texttt{record\_full\_arg} option
mean that, in general $h_{ \gamma }$ corresponds to the length of one or more
edges.
\end{itemize}

\label{app-multiple-mergers}
\section*{Multiple merger coalescent model}

Multiple merger coalescents, in which a random number of ancestral
lineages may merge into a common ancestor at a given time,
are referred to as $\Lambda$-coalescents.
The rate at which  a given group of $k$ out of a total of  $b$ lineages merges  is
\begin{equation}\label{lambdabk}
\lambda_{b, k} =  \int_0^1  x^{k-2}(1-x)^{b-k}\Lambda(dx) + a\one{k=2}, \quad 2 \le k \le b,
\end{equation}
where $a \geq 0$ is a constant and $\Lambda$ is a finite measure on the unit interval without an atom at zero \citep{DK99,P99,S99}.
There is also a larger class of simultaneous multiple merger coalescents involving
simultaneous mergers of distinct groups of lineages into several common ancestors \citep{S00}.
These are commonly referred to as $\Xi$-coalescents, and
often arise from population models incorporating diploidy or more general polyploidy
\citep{BBE13,blath2016site}.
To describe a general $\Xi$-coalescent, let $\Delta$ denote the
infinite simplex
\begin{equation*}
 \Delta := \Bigg\{ (x_1,x_2,  \ldots ): x_1 \ge x_2 \ge \cdots \ge 0, \sum_{j = 1}^{ \infty} x_j \le 1\Bigg\}.
 \end{equation*}
The rate of mergers is determined by
$\Xi = \Xi_0 + a\delta_0$, where $a \geq 0$ is a constant, $\delta_0$ is the Dirac delta
measure, and $\Xi_0$ is a finite measure on $\Delta$ with no atom at (0, 0, \ldots).
For an initial number of blocks $b \geq 2$ and
$r \in \{1,2, \ldots, b - 1 \}$, let $k_1 \geq 2, \ldots, k_r \ge 2$ be the sizes of $r$ merger events
and $s = b - k_1 - \cdots - k_r$ be the number of blocks not participating in any merger.
The rate of each possible set of mergers with sizes $(k_1, \ldots, k_r)$ is
\begin{align*}
  \lambda_{n; k_1, \ldots, k_r; s}  = {}& \int_\Delta  \sum_{\ell = 0}^s \sum_{\substack {i_1, \ldots, i_{r+\ell} = 1\\ \text{all distinct}} }^\infty  \binom{s}{\ell} x_{i_1}^{k_1} \cdots  x
_{i_{r}}^{k_r} x_{i_{r+1}} \cdots x_{i_{r+\ell}}\left(1 - \sum_{j=1}^\infty x_j \right)^{s-\ell} \frac{1}{ \sum_{j=1}^\infty x_j^2 } \Xi_0(dx)   \\
  & +  a\one{r=1, k_1 = 2},
\end{align*}
and the number of such $(k_1, \ldots, k_r)$ mergers is
\begin{equation*}
    \mathcal{N}(b; k_1, \ldots, k_r ) = \binom{b}{k_1 \ldots k_r\, s} \frac{1}{ \prod_{j=2}^b\ell_j!  },
\end{equation*}
where $\ell_j := \#\{ i \in \{ 1, \ldots, r \} : k_i = j \}$ is the number of mergers of size $j \geq 2$ \citep{S00}.

Viewing coalescent processes strictly as mathematical
objects, it is clear that the class of $\Xi$-coalescents contains
$\Lambda$-coalescents as a specific example in which at most
one group of lineages can merge at each time, and the class of
$\Lambda$-coalescents contain the Kingman-coalescent as a special
case.  However, viewed as limits of ancestral processes derived from
specific population models they are not nested. For example, one can
obtain $\Lambda$-coalescents from haploid population models
incorporating sweepstakes reproduction and high fecundity,
and $\Xi$-coalescents for the same models for diploid populations \citep{BBE13}.
One should therefore apply the models as appropriate, i.e.\
$\Lambda$-coalescents to haploid (e.g.\ mtDNA) data,
and $\Xi$-coalescents to diploid or polyploid (e.g.\ autosomal) data \citep{blath2016site}.

In \msprime\ we have incorporated two examples of multiple-merger coalescents.
One is a diploid extension \citep{BBE13} of the haploid Moran model adapted
to sweepstakes reproduction considered by \cite{EW06}.
Let $N$ denote the population size, and take $\psi \in (0,1]$ to be fixed.
In every generation, with probability $1-\varepsilon_N$ a single individual
(picked uniformly at random) perishes.
With probability $\varepsilon_N$, $\lfloor \psi N \rfloor$ individuals picked uniformly
without replacement perish instead.
In either case, a surviving individual picked uniformly at random produces enough offspring
to restore the population size back to $N$.  Taking $\varepsilon_N =
1/N^\gamma$ for some $\gamma > 0$, \cite{EW06} obtain $\Lambda$-coalescents
for which the $\Lambda$ measure in \eqref{lambdabk} is a
point mass at $\psi$.  The simplicity of this model does allow one to obtain
some explicit mathematical results (see e.g.\
\cite{Der2012, EF2018,Freund2020,Matuszewski2017}), and the model has also been
used to simulate gene genealogies within phylogenies \citep{zhu2015hybrid}.
As well as the haploid model of \cite{EW06}, \msprime\ provides the diploid version
of \cite{BBE13}, in which individuals perish as above, but replacements are
generated by sampling a single pair of diploid individuals as parents, with
children sampling one chromosome from each parent.
Hence, there are four parent chromosomes involved in each reproduction event, which can lead to
up to four simultaneous mergers, giving rise to a $\Xi$-coalescent with merger rate
\begin{esplit}{Cconst} \lambda^{\text{Dirac}}_{b; k_1, \ldots,
k_r;s } &= \frac{c \psi^2 / 4}{1 + c \psi^2 / 4} \frac{4}{ \psi^2}\sum_{\ell = 0}^{s \wedge (4-r)} \binom{s}{\ell}
(4)_{r+\ell} (1-\psi)^{s-\ell }\left( \frac{\psi}{4} \right) ^{k_1 + \cdots +
k_r + \ell} + \frac{\one{r=1, k_1 = 2}}{1 + c \psi^2 / 4},  \\ \end{esplit}
The interpretation of \eqref{Cconst} is that `small' reproduction events in which
two lineages merge occur at rate $1/(1 + c\psi^2/4)$, while large reproduction events
with the potential to result in simultaneous multiple mergers occur at rate $(c\psi^2/4) / (1 + c\psi^2/4)$.

The other multiple merger coalescent model incorporated in \msprime\ is
the haploid population model considered by \cite{schweinsberg03},
as well as its diploid extension \citep{BLS15}.  In the haploid
version, in each generation of fixed size $N$, individuals produce random numbers of
juveniles $(X_1, \ldots, X_N)$ independently, each distributed according to a stable law satisfying
\be\label{jX} \lim_{k\to \infty} C k^\alpha \prb{X \ge k} = 1 \ee
with index $\alpha > 0$, and where $C > 0$ is a normalising constant.
If the  total number of juveniles $S_N := X_1 + \ldots + X_N$ produced
in this way is at least $N$, then $N$ juveniles are sampled uniformly at random
without replacement to form the next generation.
As long as $\EE{X_1} > 1$, one can show that $\{ S_N < N \}$ has exponentially small
probability in $N$, and does not affect the resulting coalescent as $N \to \infty$ \citep{schweinsberg03}.
If $\alpha \ge 2$ the ancestral process converges to the Kingman-coalescent; if
$1 \le \alpha < 2$ the ancestral process converges to a $\Lambda$-coalescent
with $\Lambda$ in \eqref{lambdabk} given by the Beta$(2-\alpha, \alpha)$ distribution, i.e.\
\be\label{Fbeta} \Lambda(dx) = \one{0< x \le 1} \frac{1}{B(2-\alpha,\alpha)}
x^{1 - \alpha}(1-x)^{\alpha - 1}  dx, \ee where
$B(a,b) = \Gamma(a)\Gamma(b)/\Gamma(a+b)$ for $a,b > 0$ is the
beta function \citep{schweinsberg03}.
This model has been adapted to diploid populations
by \cite{BLS15}, and the resulting coalescent is $\Xi$-coalescent with merger rate
\be\label{xibeta} \lambda_{b;k_1, \ldots, k_r; s}^{\text{Beta}} = \sum_{\ell = 0}^{ s\wedge (4-r) } \binom{s}{\ell} \frac{ (4)_{r+\ell} }{4^{k+\ell}}
\frac{B(k+\ell - \alpha, s-\ell + \alpha ) }{B(2-\alpha,\alpha)}, \ee
where $k := k_1 + \ldots + k_r$ \citep{blath2016site,BLS15}. The interpretation of \eqref{xibeta} is  that the
random number of lineages participating in a potential merger is governed by the
$\Lambda$-coalescent with rate \eqref{Fbeta}, and all participating lineages are
randomly allocated into one of four groups corresponding to the four parental
chromosomes, giving rise to up to four simultaneous mergers.

The stable law \eqref{jX} assumes that individuals can produce arbitrarily
large numbers of juveniles. Since juveniles are at least fertilised eggs,
it may be desirable to suppose that the number of juveniles
surviving to reproductive maturity cannot be arbitrarily large.  Hence we also consider
an adaptation of the \cite{schweinsberg03} model, where
the random number of juveniles has a deterministic upper bound $\phi(N)$,
and the distribution of the number of juveniles produced by a given parent
(or pair of parents in the diploid case) is
\be\label{jtr} \prb{X=k} =   \one{1 \le k \le \phi(N)}
\frac{\phi(N+1)^\alpha }{ \phi(N+1)^\alpha - 1 }  \left( \frac{1}{k^\alpha} -
\frac{1}{(k+1)^\alpha}  \right) . \ee
See \citet{Eldon2018} for a related model.   One can follow the calculations of
\cite{schweinsberg03} or \cite{BLS15}  to show  that  if $1 < \alpha < 2$
then, recalling that $k = k_1 + \cdots + k_r$, the merger rate is
 \be \lambda_{b;k_1, \ldots, k_r; s}^{\text{Beta}, M} =  \sum_{\ell = 0}^{ s \wedge (4-r) }
\binom{s}{\ell} \frac{ (4)_{r+\ell} }{4^{k+\ell}} \frac{B(M; k+\ell - \alpha,
s-\ell + \alpha ) }{B(M;2-\alpha,\alpha)}
\ee
where $B(z;a,b) := \int_0^z
t^{a-1}(1-t)^{b-1}dt$ for  $a,b>0$ and $0< z\le 1$ is the incomplete beta function, and
\begin{equation*}
M :=  \lim_{ N \to \infty } \frac{\phi(N) / N}{\phi(N) / N + \EE{X_1}} \in (0, 1]
\end{equation*}
\citep{CDEE2020}.    In other words,  the measure $\Lambda$ driving
the multiple mergers is of the same form as in \eqref{Fbeta}  with $0 < x
\le M$ in the case $1 < \alpha < 2$ and $\lim_{ N \to \infty } \phi(N) / N > 0$.   If $\alpha \ge 2$
or $\phi(N)/N \to 0$ then  the  ancestral process converges to  the Kingman-coalescent
\citep{CDEE2020}.

\label{app-sweeps}
\section*{Selective sweep trajectories}
We have implemented the coalescent with selective sweeps as a structured
coalescent process (as in~\citep{braverman1995hitchhiking}). During the
``sweep phase'' of the simulation the rate of coalescence and recombination
is tracked separately in lineages linked to the beneficial allele or not,
with those rates depending on the frequency of the beneficial allele in
the population at a given time. The allele frequency trajectory of the
sweeping allele is modelled using a jump process that approximates
the conditional diffusion of the allele, conditioning on its eventual
fixation~\citep{coop2004ancestral}.
The jump process moves back in time following the beneficial
allele frequency, $p$, from some initial frequency (e.g., $p=1$) back to
the origination of the allele at $p=1/2N$, and we track time in small
increments $\delta t$. This works such that the allele frequency at
the next time step back in time, call it $p^\prime$ is

\begin{eqnarray*}
	p'=\left\{ \begin{array}{ll}
	    p + \mu(p)\delta t +\sqrt{p(1-p)\delta t} & \quad \mbox{with probability $1/2$}\\
	    p + \mu(p)\delta t -\sqrt{p(1-p)\delta t} & \quad \mbox{with probability $1/2$}\\
	\end{array} \right.
\end{eqnarray*}
where $\alpha=2Ns$ and $s$ is the fitness advantage in homozygotes. This model
assumes genic selection (i.e., that the dominance coefficient $h=0.5$). Further
$\mu(p)$ is
\begin{eqnarray*}
	\mu(p)=\frac{\alpha p (1 - p)}{tanh(\alpha(1-p))}
\end{eqnarray*}

While our current implementation only deals with genic selection,
we plan in the future to add sweeps with arbitrary dominance. Further planned
additions include trajectories of purely neutral alleles, as well as trajectories
that model soft selective sweeps.

\end{document}