### 2021-09-19 cover_letter/cover_letter.tex
Cover letter.

\documentclass{letter}

\signature{Jerome Kelleher}

\address{Big Data Institute\\University of Oxford\\UK}
\begin{document}

\begin{letter}{GENETICS}

\opening{Dear Editors,}

I am writing on behalf of my coauthors to submit our manuscript entitled
\emph{Efficient ancestry and mutation simulation with msprime 1.0}
for consideration for publication in GENETICS
as a ``Methods, Technology, and Resources'' full length research article.

Since its introduction in 2016, \texttt{msprime} has grown in popularity
and is now challenging \texttt{ms} as the widely used coalescent simulator
(according to Google Scholar, \texttt{ms} has had 100 citations so far
in 2021 and \texttt{msprime} 107). This manuscript marks the important
1.0 release milestone and represents the work
of a large community. We illustrate the large range of features
that \texttt{msprime} has accumulated since the 2016 publication,
and show that performance is excellent,
often many times faster and more memory efficient than specialised
alternative simulators.

We believe that this manuscript is a summary of the state-of-the-art
in coalescent simulation, and it will be of substantial
interest to the readers of GENETICS. We hope that you agree.

\closing{Sincerely,}

\end{letter}
\end{document}