### 2021-12-01 cover_letter/cover_letter2.tex

\documentclass{letter}

\signature{Jerome Kelleher}

\address{Big Data Institute\\University of Oxford\\UK}
\begin{document}

\begin{letter}{GENETICS}

\opening{Dear Editors,}

I am writing on behalf of my coauthors to resubmit our manuscript entitled
\emph{Efficient ancestry and mutation simulation with msprime 1.0}
for consideration for publication in GENETICS
as a ``Methods, Technology, and Resources'' full length research article.

We thank the reviewers for their thoughtful and thorough reviews. We
believe that we have answered all questions raised, leading to a more
concise and consistently presented manuscript. We have attached
a detailed point-by-point response in the \texttt{response-to-reviewers.pdf}
file.

We have also attached a file showing the changes made to the manuscript
in response to reviewers in \texttt{review-diff.pdf}.

Finally, we attach \texttt{ms-genetics-format.pdf}, which is the manuscript
reformatted using the GENETICS latex template, and converted to US English.

\closing{Sincerely,}

\end{letter}
\end{document}