### 2021-12-06 cover_letter/cover_letter3.tex
Cover letter for final submission

\documentclass{letter}

\signature{Jerome Kelleher}

\address{Big Data Institute\\University of Oxford\\UK}
\begin{document}

\begin{letter}{GENETICS}

\opening{Dear Editors,}

Please see enclosed the final version of our manuscript in PDF format.
I am unclear what the requirements for the Latex source and image
files are, and will therefore await instructions.

Unfortunately, I will need to revise the grant data provided in the
online form, as it is not complete. Can you advise on how to do this please?

\closing{Sincerely,}

\end{letter}
\end{document}